### 2023-10-17 scripts/seg_sites_rarefaction.Rmd
added scripts

---
title: "Seg_sites_rarefaction"
output: html_document
date: "2023-08-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table)

```

# Generate data
```{r}
# Generate some example data:
## define parameters
allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
n1 <- 2000 # p1 gene copy number
n2 <- 200 # p2 gene copy number
k <- 30 # number to rarefact to
mcmcs <- 10000 # number of mcmc draws

## generate genotypes
g1 <- matrix(0, nrow = length(allele_frequencies), ncol = n1) # init
g2 <- matrix(0, nrow = length(allele_frequencies), ncol = n2)

m1 <- lapply(n1*allele_frequencies, function(x) sample(n1, x, FALSE))  # which are the minor
m2 <- lapply(n2*allele_frequencies, function(x) sample(n2, x, FALSE))

populate <- function(g, minors){g[minors] <- 1; return(g)} # function to replace minors
for(i in 1:length(allele_frequencies)){ # do for every locus
  g1[i,] <- populate(g1[i,], m1[[i]])
  g2[i,] <- populate(g2[i,], m2[[i]])
}

## combine into individuals
g1 <- g1[,1:(ncol(g1)/2)] + g1[,((ncol(g1)/2) + 1):ncol(g1)]
g2 <- g2[,1:(ncol(g2)/2)] + g2[,((ncol(g2)/2) + 1):ncol(g2)]
```

# Via rarefaction approach
```{r}
counts <- function(g){
  return(data.frame(majhom = rowSums(g == 0), het = rowSums(g == 1), minhom = rowSums(g == 2)))
}
rarefact <- function(counts, k){
  return(1 - ((choose(counts[,1], k) + choose(counts[,3], k))/choose(rowSums(counts), k)))
}
varseg <- function(p) p*(1-p)

# taken from the binom package
confint_pb <- function(p, n, alpha){
  alpha <- rep(alpha, length = length(p))
  alpha2 <- 0.5 * alpha
  z <- qnorm(1 - alpha2)
  z2 <- z * z
  se <- sqrt(p * (1 - p)/n)
  lcl <- p - z * se
  ucl <- p + z * se

  lcl[lcl < 0] <- 0
  ucl[ucl > 1] <- 1
  return(data.frame(lower = lcl, upper = ucl))
}

# confint_sum <- function(nT, n, alpha){
#   Tbar <- nT/n
#   
#   if(n*Tbar == 0){return(0)}
#   
#   alpha_star <- zipfR::Rbeta((n*Tbar - 1)/n, n*Tbar, (n*(1 - Tbar)) + 1)
#   
#   if(alpha_star <= alpha & alpha <= 1){
#     top <- 1 - alpha
#     bottom <- n*(1 - alpha_star)
#     res <- Tbar - (top/bottom)
#   }
#   if(0 <= alpha & alpha <= alpha_star){
#     return(zipfR::Rbeta.inv(alpha, n*Tbar, n*(1 - Tbar) + 1))
#   }
#   
# }

c1 <- counts(g1) # geno counts
c2 <- counts(g2)

pseg1 <- rarefact(c1, k) # rarefact
pseg2 <- rarefact(c2, k) # rarefact

vseg1 <- varseg(pseg1)
vseg2 <- varseg(pseg2)
ciseg1 <- confint_pb(pseg1, k, .05)
ciseg2 <- confint_pb(pseg2, k, .05)

p1 <- sum(pseg1)/length(allele_frequencies)
p2 <- sum(pseg2)/length(allele_frequencies)
```

# Via mcmc

```{r}
rarefact_mcmc <- function(g, k){
  samp <- sample(g, k, FALSE)
  seg <- ifelse(any(samp == 1), 1, # got hets
                ifelse(any(samp == 0) & any(samp == 2), 1, 0)) # got both homs
  return(seg)
}

# draw
seg1 <- seg2 <- matrix(0, length(allele_frequencies), mcmcs)
for(i in 1:mcmcs){
  for(j in 1:length(allele_frequencies)){
    seg1[j,i] <- rarefact_mcmc(g1[j,], k = k)
    seg2[j,i] <- rarefact_mcmc(g2[j,], k = k)
  }
}

mpseg1 <- binom::binom.confint(rowSums(seg1), rep(mcmcs, nrow(seg1)), 
                               method = "exact")[,c("mean", "lower", "upper")]
mpseg2 <- binom::binom.confint(rowSums(seg2), rep(mcmcs, nrow(seg1)),
                               method = "exact")[,c("mean", "lower", "upper")]

colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")

```

```{r}
# mcmcs <- 1000
# res <- d@geno.tables$gs
# res <- data.table::as.data.table(res)
# 
# matches <- which(d@facet.meta$facet == "pop")
# res <- res[matches,]
# 
# gs <- d@geno.tables$gs[matches,]
# gs <- data.table::as.data.table(gs)
# gs$.sum <- rowSums(gs) # sums for each row
# gs <- cbind(as.data.table(d@facet.meta[matches,]), gs)
# gs[,.g := min(.sum) - 1, by = .(facet, .snp.id)] # min across all levels
# 
# fwrite(res, "test_gs.txt", col.names = TRUE, sep = "\t", eol = "\n")
# fwrite(data.table(g = gs$.g), "test_g.txt", col.names = FALSE, sep = "\t", eol = "\n")
# fwrite(data.table(g = gs$subfacet), "test_subfacets.txt", col.names = FALSE, sep = "\t", eol = "\n")
# 
# 
# cmd <- paste0("perl seg_sites_rarefaction.pl test_gs.txt test_g.txt test_subfacets.txt ", mcmcs, " test_out.txt")
# system(cmd)
# 
# mcmc_res <- fread("test_out.txt")
# fc <- ncol(mcmc_res)
# mcmc_res <- mcmc_res[,-..fc]
# colnames(mcmc_res)[1] <- "subfacet"
# mcmc_res <- melt(mcmc_res, id.vars = "subfacet")
# 
# tr <- function(x){res <- t.test(x); return(list(seg_sites = res$estimate, lowerCI = res$conf.int[1], upperCI = res$conf.int[2]))}
# mcmc_summary <- mcmc_res[,tr(value), by = subfacet]

```

# results
```{r}
# individual loci
pseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = pseg1, snp = 1:length(allele_frequencies), af = allele_frequencies),
              data.frame(pop = paste0("n = ", n2/2), p = pseg2, snp = 1:length(allele_frequencies), af = allele_frequencies))

oseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = mpseg1[,1], snp = 1:length(allele_frequencies), af = allele_frequencies),
              data.frame(pop = paste0("n = ", n2/2), p = mpseg2[,1], snp = 1:length(allele_frequencies), af = allele_frequencies))

pseg$source <- "Mathematical"
oseg$source <- "Simulation"

seg <- rbind(pseg, oseg)
cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("n = ", n1/2), af = allele_frequencies),
             cbind.data.frame(mpseg2, pop = paste0("n = ", n2/2), af = allele_frequencies))
colnames(cis) <- c("p", "lower", "upper", "pop", "af")
seg <- merge(seg, cis[,c("lower", "upper", "af", "pop")], by = c("af", "pop"))
seg$in_ci <- ifelse(seg$p >= seg$lower & seg$p <= seg$upper, TRUE, FALSE)

ggplot(seg, aes(x = af, y = p, color = source)) + geom_point() + facet_wrap(~pop)

ggplot(seg[seg$source == "Mathematical",],
       aes(x = af, y = p)) +
  geom_point(aes(color = in_ci)) +
  geom_errorbar(data = cis, aes(ymax = upper, ymin = lower)) +
  facet_wrap(~pop) +
  theme_bw()

# sums across loci
msum1 <- colSums(seg1)
msum2 <- colSums(seg2)

msum_tab <- data.frame(nseg = c(mean(msum1), mean(msum2)),
                       lower = c(quantile(msum1, .05), quantile(msum2, .05)),
                       upper = c(quantile(msum1, .95), quantile(msum2, .95)),
                       pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)))

sum_tab <- tapply(seg[seg$source == "Mathematical",]$p, seg[seg$source == "Mathematical",]$pop, sum)
sum_tab <- as.data.frame(sum_tab)
sum_tab$pop <- rownames(sum_tab)
colnames(sum_tab)[1] <- "nseg"

all_mcmc_sums <- rbind(cbind.data.frame(nseg = msum1, pop = paste0("n = ", n1/2)),
                   cbind.data.frame(nseg = msum2, pop = paste0("n = ", n2/2)))

ggplot(sum_tab, aes(x = pop, y = nseg)) + 
  geom_point(color = "red") +
  geom_errorbar(data = msum_tab, aes(ymax = upper, ymin = lower, x= pop), inherit.aes = FALSE) +
  # geom_point(data = msum_tab, color = "blue") +
  geom_jitter(data = all_mcmc_sums, height = 0, alpha = .005) +
  theme_bw() +
  scale_y_continuous(limits = c(0, length(allele_frequencies)))

tapply(seg[seg$source == "Mathematical",]$in_ci, seg[seg$source == "Mathematical",]$pop, mean, na.rm = TRUE)

# variances, per locus
vars <- rbind(cbind.data.frame(estimated = vseg1, 
                               observed = matrixStats::rowVars(seg1),
                               pop = paste0("n = ", n1/2),
                               af = allele_frequencies),
              cbind.data.frame(estimated = vseg2, 
                               observed = matrixStats::rowVars(seg2),
                               pop = paste0("n = ", n2/2),
                               af = allele_frequencies))
ggplot(vars, aes(x = observed, y = estimated, color = af)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  facet_wrap(~pop) +
  theme_bw() +
  khroma::scale_color_batlow()

# variances, overall
overall_vars <- rbind(cbind.data.frame(estimated = sum(vseg1),
                                       observed = var(msum1),
                                       pop = paste0("n = ", n1/2)),
                      cbind.data.frame(estimated = sum(vseg2),
                                       observed = var(msum2),
                                       pop = paste0("n = ", n2/2)))

# estimated PI -- normal approx works well
pi1 <- qnorm(.975)*sqrt(sum(vseg1))*sqrt(1 + (1/length(allele_frequencies)))
pi2 <- qnorm(.975)*sqrt(sum(vseg2))*sqrt(1 + (1/length(allele_frequencies)))
pi <- data.frame(pop = paste0("n = ", n1/2), 
                 lower = sum(pseg1) - pi1,
                 upper = sum(pseg1) + pi1)
pi <- rbind(pi, data.frame(pop = paste0("n = ", n2/2), 
                 lower = sum(pseg2) - pi2,
                 upper = sum(pseg2) + pi2))

all_mcmc_sums <- as.data.table(all_mcmc_sums)
all_mcmc_sums[pop == pi$pop[1], 
              in_pi := nseg <= pi$upper[1] & 
                nseg >= pi$lower[1]]
all_mcmc_sums[pop == pi$pop[2], 
              in_pi := nseg <= pi$upper[2] & 
                nseg >= pi$lower[2]]

ggplot(all_mcmc_sums, aes(x = pop, y = nseg, color = in_pi)) +
  geom_jitter(height = 0, alpha = .5) +
  geom_errorbar(data = pi, aes(ymax = upper, ymin = lower, x = pop), inherit.aes = FALSE)

tapply(all_mcmc_sums$in_pi, all_mcmc_sums$pop, mean)
```

### 2023-10-17 scripts/seg_sites_rarefaction.pl

@@ -1,0 +1,152 @@
+use List::Util qw/sum/;
+use List::Util qw/shuffle/;
+use Data::Dumper;
+
+if ($#ARGV == 4) {
+  $file = $ARGV[0];
+  $file2 = $ARGV[1];
+  $file3 = $ARGV[2];
+  $mcmcs = $ARGV[3];
+  $outfile = $ARGV[4];
+} 
+else {
+  die;
+}
+
+open(FILE, "<$file") or die $!;
+open(FILE2, "<$file2") or die $!;
+open(FILE3, "<$file3") or die $!;
+open(OUTFILE, '>', $outfile) or die $!;
+
+my @output;
+
+$header = <FILE>;
+chomp($header);
+@header = split(/\t/, $header);
+
+my @stat = stat $file;
+
+
+
+# read in the subfacet details
+my @subfacets;
+my $tracker = 0;
+while(<FILE3>){
+  chomp($_);
+  $pop = $_;
+  $pop =~ s/\R//g;
+  $subfacets[$tracker] = $pop;
+  $tracker++;
+}
+
+close FILE3;
+
+# init results
+my %results;
+@pops = uniq(@subfacets);
+foreach $pop (@pops){
+  @{$results{$pop}} = (0) x $mcmcs;
+}
+
+my $locus_tracker = 0;
+
+print("Starting...\n");
+while (<FILE>) {
+  if($locus_tracker % 10000 == 0){
+      print("Loci: ", $locus_tracker, "\n");
+  }
+  
+  #print("=============================================\n");
+  
+  $genotypes = $_;
+  chomp($genotypes);
+  
+  $g = <FILE2>;
+  $g =~ s/\R//g;
+  next if($g < 1);
+  #print("g = ", $g, "\n");
+
+  
+
+  @genotypes = split(/\t/, $genotypes);
+  
+  $n = sum(@genotypes);
+  
+  # figure out our population
+  $tpop = $subfacets[$locus_tracker];
+  
+  my @genoarray = ();
+  
+  # build sampling array
+  $i = 0;
+  for($i; $i <= $#genotypes; $i++){
+    if($genotypes[$i] != 0){
+      #print($i, " filling with ", $genotypes[$i], " ", $header[$i], "'s\n");
+      @narray = ($header[$i]) x $genotypes[$i];
+      push(@genoarray, @narray);
+    }
+  }
+  
+  #foreach(@genoarray){print($_, " ");}
[…]

### 2023-10-17 manuscript/ms.Rmd
started work on paper

---
title: "manuscript"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    reference_docx: "paper-files/style_ref.docx"
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
  html_document:
    df_print: paged
bibliography: "paper-files/citations.bib"
csl: "paper-files/apa.csl"
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{setspace}\doublespacing
- \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
- \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
- \usepackage[round]{natbib}
indent: yes
sansfont: Times New Roman
fontsize: 12pt
always_allow_html: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(snpR); library(data.table); library(ggplot2); library(cowplot)
```

# Abstract

# Introduction
<!-- Outline: -->

<!-- * What is genetic diveristy? -->
<!--   - Utility -->
<!--   - Estimators -->
<!--     - heterozygosity -->
<!--     - pi -->
<!--     - private alleles -->
<!--     - allele counts per locus -->

<!-- * Problems with ac and private alleles per locus -->
<!--   - All subject to bias due to differences in effective sample size -->
<!--     - Missing data, unequal sample sizes -->

<!-- * Rarefaction -->
<!--   - Sampling vs mathematical solutions -->
<!--   - solutions for private alleles and allele counts (richness) -->
<!--   - description -->
<!--   - how this solves problems -->

<!-- * Nseg -->
<!--   - Newer metric for modern sequencing tech -->
<!--   - Useful alternative, lets you capture different parts of variance -->
<!--   - biases -->
<!--   - rarefaction as solution -->

# Methods
Outline:

* Background on methods for private alleles, richness
* Solution for segregating sites
  - Derivation
  - Can account for both different sample sizes and missingness across loci and populations.
* Simulations
  - Sampling approach, missingness, etc

<!-- https://link.springer.com/article/10.1023/B:COGE.0000041021.91777.1a#citeas -->

## Probability of observing segregating loci via rarefaction
Rarefaction can be used to estimate the probability of observing a segregating site at a specific locus using much the same framework used to calculate allelic richness. In brief, allelic richness (or the expected number of distinct alleles expected at a given locus under a common sample size $g$ across populations) can be estimated for a given population $j$ by summing the probability of observing each $i$ of $m$ unique alleles using the counts of those alleles in the population $N_{ij}$ and the total sample size in that population $N_{j}$. This is done by comparing the number of possible ways to draw $g$ gene copies without sampling allele $i$ ($\binom{N_{j} - N_{ij}}{g}$) to the total number of possible combinations of gene copies that can be drawn ($\binom{N_{j}}{g}$); the inverse of this ($1 - \frac{\binom{N_{j} - N_{ij}}{g}}{\binom{N_{j}}{g}}$) is therefore the probability of observing allele $i$ in population $N_{j}$, and the sum of this value across all $m$ alleles gives the expected number of alleles observed at a locus in population $j$, $\alpha_{g}^{j}$ [@hurlbertNonconceptSpeciesDiversity1971;@kalinowskiCountingAllelesRarefaction2004]:
$$\alpha_{g}^{j} = \sum_{i = 1}^{m} 1 - \frac{\binom{N_{j} - N_{ij}}{g}}{\binom{N_{j}}{g}}$$

The expected number of segregating loci in a population for a draw of $g$ gene copies can be derived similarly. For a locus $i$ to be segregating in population $j$, all alleles drawn across all gene copies must be identical. If alleles are independent at each locus (the locus is at Hardy-Weinburg Equlibrium, HWE) and $N$ is infinite, the probability ($P({S_{j}})$) of observing a segregating site at a locus is the inverse of the probability of drawing only one allele in $g$ draws with replacement: $$P({S_{j}}) = 1 - \sum_{i = 1}^{m} f_{ij}^{g}$$

where $f_{ij}$ is the allele frequency of allele $i$ in population $j$. However, in finite samples draws are conducted with replacement, and so binomial coefficients must instead be used to determine the probability of drawing only a specific allele:
$$P({S_{j}}) = 1 - \sum_{i = 1}^{m} \frac{\binom{N_{ij}}{g}}{\binom{N_{j}}{g}}$$

However, HWE is often not a desirable assumption to make. Even if filtering is employed to remove loci which do not conform to HWE, the degree of conformity, and thus the degree of statistical bias in estimating $P({S_{j}})$, typically varies somewhat between populations.

To remedy this, I propose the following estimator of $P({S_{j}})$:
$$P({S_{j}}) = 1 - \sum_{k = 1}^{h}\frac{\binom{n_{kj}}{\gamma}}{\binom{n_{j}}{\gamma}}$$
where the $P({S_{j}})$ is given by the probability of exclusively drawing any $k$ of $h$ possible homozygote genotypes in population $j$ given $\gamma$ independent sampled *genotypes* (not *gene copies*) from the pool of observed genotypes. Here, $n_{kj}$ is the number of observed $k$ genotypes in population $j$ where the total number of observed genotypes of all types is $n_{j}$. Note that $n_{j}$ and $\gamma$ will be half the value of their equivalents $N_{j}$ and $g$ for diploid species, one third for triploids, and so on.

Interestingly, this method, like the richness method and related private allele rarefaction approaches can smoothly account for varying amounts of missing data at specific loci in different populations by varying $g$ or $\gamma$ across loci. Both can be set to one less than the smallest observed $N_{j}$ or $n_{j}$, the highest values at which rarefaction can be applied within a population, across all populations after accounting for missing data, and can thus vary across loci without bias. Setting either value to $N_{j}$ or $n_{j}$ will instead return the observed allele diversity or segregating site status, respectively.

This is particularly useful given that $E(N_{S})$ or the expected total number of segregating sites, is often of specific interest as a measure of genetic diversity when comparing populations. Given that the expected number of segregating sites at a specific locus $q$ in population $j$, $E(N_{S_{jq}})$, is equal to $P({S_{jq}})$, $E(N_{S})$ can be calculated by summing $P({S_{jq}})$ across all $Q$ loci: $$E(N_{S}) = \sum_{q = 1}^{Q} P({S_{jq}})$$
with $\gamma$ set accordingly for each locus. In this case, $0 \le E(N_{S_{jq}}) \le 1$ for all loci (and thus $0 \le E(N_{S}) \le Q$).

Usefully, under this framework each locus represents a single Bernoulli trial in which it can be observed to be segregating or not with probability $P({S_{jq}})$. As such, the variance of $P({S_{jq}})$ for each locus is given by $$\sigma_{P(S_{jq})}^{2} = P(S_{jq}) \times (1 - P(S_{jq}))$$
and, if each locus is independent, the variance of $E(N_{S})$ is equal to the sum of $\sigma_{P(S_{jq})}^{2}$ across all loci:
$$\sigma_{E(n_{S})}^{2} = \sum_{q = 1}^{Q}\sigma_{P(S_{jq})}^{2}$$

Confidence and prediction intervals can then be derived using standard approaches for the sum of random, independent Bernoulli trials. When $Q$ is large, for example, the distribution of $E(N_{S})$ should be approach normal and confidence and prediction intervals can be derived using standard normal approximation. In contrast, confidence intervals for individual $P({S_{jq}})$ values can be achieved via standard approaches for single Bernoulli samples of size $\gamma$.

## Emperical Validation
To validate equations 4 and 5, I simulated genotypic data for two populations with sizes 100 and 1000, each with 100 bi-allelic loci with minor allele frequencies spaced equally between 0.01 and 0.1. I added missing data to each population assigning each locus a missing data rate $R_{mq}$ from a uniform distribution such that $R_{mq} \sim U(0,.3)$, ensuring that overall allele frequencies in each population were maintained. I then used the methods described above to estimate $P({S_{jq}})$ and $E(N_{S})$ and their variances given $\gamma = 30$. I used the "asymptotic" method from the R package `binom` [@rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each $P({S_{jq}})$ and normal approximation to calculate 95% confidence and prediction intervals for $E(N_{S})$.

For comparison, I also conducted 10,000 random draws of size $\gamma$ from each loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws. I likewise calculated the variance of $E(N_{S})$ directly across all sets of random draws.

# Results
The methods described here for calculating $P({S_{j}})$ and $E(N_{S})$ performed well. 95% confidence intervals for simulated loci overlapped the simulated mean in

### 2023-10-17 scripts/seg_sites_rarefaction.Rmd
started work on paper

@@
 ```
-
-# Generate data
+# Define parameters
 ```{r}
@@
 ## define parameters
+seed <- 231232
 allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
@@
 mcmcs <- 10000 # number of mcmc draws
+max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
 
-## generate genotypes
-g1 <- matrix(0, nrow = length(allele_frequencies), ncol = n1) # init
-g2 <- matrix(0, nrow = length(allele_frequencies), ncol = n2)
+## derived
+set.seed(seed)
+nal <- length(allele_frequencies)
+miss_percent <- floor(runif(nal, 0, max_missing))/100
 
-m1 <- lapply(n1*allele_frequencies, function(x) sample(n1, x, FALSE))  # which are the minor
-m2 <- lapply(n2*allele_frequencies, function(x) sample(n2, x, FALSE))
+```
 
-populate <- function(g, minors){g[minors] <- 1; return(g)} # function to replace minors
-for(i in 1:length(allele_frequencies)){ # do for every locus
-  g1[i,] <- populate(g1[i,], m1[[i]])
-  g2[i,] <- populate(g2[i,], m2[[i]])
+# Generate data
+```{r}
+# generate genotypes
+g1 <- matrix(0, nrow = nal, ncol = n1) # init
+g2 <- matrix(0, nrow = nal, ncol = n2)
+
+# add missing data
+ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
+ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
+
+# determine minor allele placements
+ma1 <- ma2 <- vector("list", length(ms1))
+for(i in 1:nal){
+  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
+  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)
+
+}
+
+# populate
+## function to replace minors
+populate <- function(g, missing, minors){
+  g[missing] <- NA
+  g[-missing][minors] <- 1
+  return(g)
+}
+## do for every locus
+for(i in 1:nal){ 
+  g1[i,] <- populate(g1[i,], ms1[[i]], ma1[[i]])
+  g2[i,] <- populate(g2[i,], ms2[[i]], ma2[[i]])
 }
@@
 counts <- function(g){
-  return(data.frame(majhom = rowSums(g == 0), het = rowSums(g == 1), minhom = rowSums(g == 2)))
+  return(data.frame(majhom = rowSums(g == 0, na.rm = TRUE), 
+                    het = rowSums(g == 1, na.rm = TRUE), 
+                    minhom = rowSums(g == 2, na.rm = TRUE)))
 }
@@
 
-# taken from the binom package
+# taken from the binom package, asymptotic method
 confint_pb <- function(p, n, alpha){
@@
 
-# confint_sum <- function(nT, n, alpha){
-#   Tbar <- nT/n
-#   
-#   if(n*Tbar == 0){return(0)}
-#   
-#   alpha_star <- zipfR::Rbeta((n*Tbar - 1)/n, n*Tbar, (n*(1 - Tbar)) + 1)
-#   
-#   if(alpha_star <= alpha & alpha <= 1){
-#     top <- 1 - alpha
-#     bottom <- n*(1 - alpha_star)
-#     res <- Tbar - (top/bottom)
-#   }
-#   if(0 <= alpha & alpha <= alpha_star){
-#     return(zipfR::Rbeta.inv(alpha, n*Tbar, n*(1 - Tbar) + 1))
-#   }
-#   
-# }
-
 c1 <- counts(g1) # geno counts
@@
 
-p1 <- sum(pseg1)/length(allele_frequencies)
-p2 <- sum(pseg2)/length(allele_frequencies)
+p1 <- sum(pseg1)/nal
+p2 <- sum(pseg2)/nal
 ```
@@
 rarefact_mcmc <- function(g, k){
+  miss <- is.na(g)
+  if(any(miss)){
+    g <- g[-which(miss)]
+  }
+
   samp <- sample(g, k, FALSE)
@@
 # draw
-seg1 <- seg2 <- matrix(0, length(allele_frequencies), mcmcs)
+seg1 <- seg2 <- matrix(0, nal, mcmcs)
 for(i in 1:mcmcs){
-  for(j in 1:length(allele_frequencies)){
+  for(j in 1:nal){
     seg1[j,i] <- rarefact_mcmc(g1[j,], k = k)
@@
 
-```{r}
-# mcmcs <- 1000
-# res <- d@geno.tables$gs
-# res <- data.table::as.data.table(res)
-# 
-# matches <- which(d@facet.meta$facet == "pop")
-# res <- res[matches,]
-# 
-# gs <- d@geno.tables$gs[matches,]
-# gs <- data.table::as.data.table(gs)
-# gs$.sum <- rowSums(gs) # sums for each row
-# gs <- cbind(as.data.table(d@facet.meta[matches,]), gs)
-# gs[,.g := min(.sum) - 1, by = .(facet, .snp.id)] # min across all levels
-# 
-# fwrite(res, "test_gs.txt", col.names = TRUE, sep = "\t", eol = "\n")
-# fwrite(data.table(g = gs$.g), "test_g.txt", col.names = FALSE, sep = "\t", eol = "\n")
-# fwrite(data.table(g = gs$subfacet), "test_subfacets.txt", col.names = FALSE, sep = "\t", eol = "\n")
-# 
-# 
-# cmd <- paste0("perl seg_sites_rarefaction.pl test_gs.txt test_g.txt test_subfacets.txt ", mcmcs, " test_out.txt")
-# system(cmd)
-# 
-# mcmc_res <- fread("test_out.txt")
-# fc <- ncol(mcmc_res)
-# mcmc_res <- mcmc_res[,-..fc]
-# colnames(mcmc_res)[1] <- "subfacet"
-# mcmc_res <- melt(mcmc_res, id.vars = "subfacet")
-# 
-# tr <- function(x){res <- t.test(x); return(list(seg_sites = res$estimate, lowerCI = res$conf.int[1], upperCI = res$conf.int[2]))}
-# mcmc_summary <- mcmc_res[,tr(value), by = subfacet]
-
-```
-
 # results
@@
 # individual loci
-pseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = pseg1, snp = 1:length(allele_frequencies), af = allele_frequencies),
-              data.frame(pop = paste0("n = ", n2/2), p = pseg2, snp = 1:length(allele_frequencies), af = allele_frequencies))
+pseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
+              data.frame(pop = paste0("n = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))
 
-oseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = mpseg1[,1], snp = 1:length(allele_frequencies), af = allele_frequencies),
-              data.frame(pop = paste0("n = ", n2/2), p = mpseg2[,1], snp = 1:length(allele_frequencies), af = allele_frequencies))
+oseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies),
+              data.frame(pop = paste0("n = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies))
 
@@
 seg$in_ci <- ifelse(seg$p >= seg$lower & seg$p <= seg$upper, TRUE, FALSE)
-
-ggplot(seg, aes(x = af, y = p, color = source)) + geom_point() + facet_wrap(~pop)
 
@@
   geom_errorbar(data = msum_tab, aes(ymax = upper, ymin = lower, x= pop), inherit.aes = FALSE) +
-  # geom_point(data = msum_tab, color = "blue") +
   geom_jitter(data = all_mcmc_sums, height = 0, alpha = .005) +
   theme_bw() +
-  scale_y_continuous(limits = c(0, length(allele_frequencies)))
+  scale_y_continuous(limits = c(0, nal))
 
@@
 # estimated PI -- normal approx works well
-pi1 <- qnorm(.975)*sqrt(sum(vseg1))*sqrt(1 + (1/length(allele_frequencies)))
-pi2 <- qnorm(.975)*sqrt(sum(vseg2))*sqrt(1 + (1/length(allele_frequencies)))
+pi1 <- qnorm(.975)*sqrt(sum(vseg1))*sqrt(1 + (1/nal))
+pi2 <- qnorm(.975)*sqrt(sum(vseg2))*sqrt(1 + (1/nal))
 pi <- data.frame(pop = paste0("n = ", n1/2), 

### 2023-12-20 manuscript/ms.Rmd
Introduction draft added

@@
 # Introduction
+
+Estimating the degree of genetic diversity has long been a critically important and widespread practice in conservation genetics [@dewoodyLongstandingSignificanceGenetic2021].
+Genetic diversity, or the diversity alleles, genotypes, and segregating loci within a population [@frankhamIntroductionConservationGenetics2002], is the both the basic foundation upon which selection acts and thus in part controls population adaptability [@Kardos2021], and is strongly correlated with fitness both at the individual and population level, even when no obvious functional ties are known [reviewed in @dewoodyLongstandingSignificanceGenetic2021].
+
+Estimators of population genetic diversity are themselves extremely diverse, but can essentially be split into two categories depending on if they are allele frequency dependent or independent.
+Allele frequency dependent statistics, such as expected heterozygosity ($H_{e}$), the F-statistics ($F_{IS}$, $F_{ST}$, and $F_{IT}$)[@Weir1984], nucleotide diversity ($\pi$), and Tajima's $\theta$ [@Tajima1989] all rely principally on the frequency of the alleles across any surveyed loci; whereas allele frequency independent statistics, such as Watterson's $\theta$ [@wattersonNumberSegregatingSites1975], counts of alleles per loci, the number of observed segregating sites, and (to some degree) private allele counts do not.
+This is not a trivial difference:  evolutionary forces do not act equally on the number of loci/alleles in a population and the frequency of those alleles.
+Tajima's D, for example, is a powerful and broadly used statistic fundamentally based on the difference between Tajima's $\theta$ and Watterson's $\theta$ can be used to both detect selection and population demographic changes [@Tajima1989] due to the way that those forces act on the balance of allele frequencies within populations.
+
+The average allele count per locus (which is usually corrected to allelic richness [@kalinowskiCountingAllelesRarefaction2004]) has historically been one of the most prevalent allele frequency independent measures of genetic diversity.
+While this is a particularly useful measure in microsatellite and other datasets utilizing heavily poly-allelic markers, it is less so in single-nucleotide polymorphism (SNP) datasets.
+SNP datasets are usually primarily biallelic, and thus the count of the number of alleles per locus in such datasets is less variable across populations and therefore less interesting.
+Counts of the number of segregating sites per population are a useful alternative in SNP datasets given the large number of possible segregating sites that these datasets often query [@?CITE].
+
+However, raw counts of the number of segregating sites per population are problematic whenever sample sizes are not equal across the populations under comparison at all loci.
+This can occur either directly due to unequal numbers of individuals or  indirectly due to unqual proportions of missing data across populations.
+This problem also effects estimates of allele counts per locus and private allele numbers, but corrections for both which use rarefaction to estimate those parameters under a common sample size are well developed [@kalinowskiCountingAllelesRarefaction2004].
+
+Here we present rarefaction-corrected estimators for the probability that any given loci would be observed to be segregating under a reduced sample size and the expected total number of segregating sites across all loci within a population.
+We show that these estimators are highly accurate via comparison to simulated re-sampling using simulated data.
+These estimators are currently implemented and available in the `snpR` R package via the function `calc_seg_sites` [@hemstromSnpRUserFriendly2023], where they can be automatically run for all populations either with automatic, per-locus optimized control of target sample size or globally via manual control.
+
 <!-- Outline: -->
@@
 # Methods
-Outline:
-
-* Background on methods for private alleles, richness
-* Solution for segregating sites
-  - Derivation
-  - Can account for both different sample sizes and missingness across loci and populations.
-* Simulations
-  - Sampling approach, missingness, etc
-
 <!-- https://link.springer.com/article/10.1023/B:COGE.0000041021.91777.1a#citeas -->
@@
 Even if filtering is employed to remove loci which do not conform to HWE, the degree of conformity, and thus the degree of statistical bias in estimating $P({S_{j}})$, typically varies somewhat between populations.
+For example, in a sample of 100 genotypes with a minor allele frequency of 0.05, only five minor alleles are expected and two out of three possible combinations of minor homozygotes and heterozygotes that produce that frequency will not deviate from HWE at $\alpha = 0.05$ according to an exact test [@Wigginton2005].
+However, re-sampling these to, say, ten genoytpes should will produce quite different $P(S_{j})$ (roughly 0.7, 0.6, and 0.5, for purely heterozygotes, one homozygote and three heterozygotes, and two homozygotes and one heterozygote, respectively)  as we will see below.
 
-To remedy this, I propose the following estimator of $P({S_{j}})$:
+To remedy these problems, I propose the following estimator of $P({S_{j}})$:
 $$P({S_{j}}) = 1 - \sum_{k = 1}^{h}\frac{\binom{n_{kj}}{\gamma}}{\binom{n_{j}}{\gamma}}$$
 where the $P({S_{j}})$ is given by the probability of exclusively drawing any $k$ of $h$ possible homozygote genotypes in population $j$ given $\gamma$ independent sampled *genotypes* (not *gene copies*) from the pool of observed genotypes.
-Here, $n_{kj}$ is the number of observed $k$ genotypes in population $j$ where the total number of observed genotypes of all types is $n_{j}$.
+Here, $n_{kj}$ is the number of observed homozygote genotypes of type $k$ in population $j$ and $n_{j}$ is the total number of observed genotypes of all types.
 Note that $n_{j}$ and $\gamma$ will be half the value of their equivalents $N_{j}$ and $g$ for diploid species, one third for triploids, and so on.
@@
 Confidence and prediction intervals can then be derived using standard approaches for the sum of random, independent Bernoulli trials.
-When $Q$ is large, for example, the distribution of $E(N_{S})$ should be approach normal and confidence and prediction intervals can be derived using standard normal approximation.
-In contrast, confidence intervals for individual $P({S_{jq}})$ values can be achieved via standard approaches for single Bernoulli samples of size $\gamma$.
+When $Q$ is large, for example, the distribution of $E(N_{S})$ should be approach normal and confidence and prediction intervals can be derived using standard normal approximation using the equations $$CI_{N_{S}} = E(N_{S}) \pm Z\sqrt{\frac{\sigma_{E(n_{S})}^{2}}{Q}}$$ $$PI_{N_{S}} = E(N_{S}) \pm Z\sqrt{\sigma_{E(n_{S})}^{2}(1+(1/Q))}$$
+where $Z$ is given by the normal quantile function $Z = Q_{X}(1-\alpha)$ with $\mu = 0$ and $\sigma = 1$ for a desired confidence level $\alpha$.
 
@@
 I then used the methods described above to estimate $P({S_{jq}})$ and $E(N_{S})$ and their variances given $\gamma = 30$.
-I used the "asymptotic" method from the R package `binom` [@rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each $P({S_{jq}})$ and normal approximation to calculate 95% confidence and prediction intervals for $E(N_{S})$.
+For comparison, I also conducted 10,000 random draws of size $\gamma$ from each loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws.
+I likewise calculated the variance of $N_{S}$ directly across all sets of random draws.
 
-For comparison, I also conducted 10,000 random draws of size $\gamma$ from each loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws.
-I likewise calculated the variance of $E(N_{S})$ directly across all sets of random draws.
+To compare my calculated estimates to the empirical simulations, I used the implementation of the Agresti-Coull [@agrestiApproximateBetterExact1998] method from the R package `binom` [@dorai-rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each empirical $P({S_{jq}})$.
+I also used normal approximation to calculate 95% confidence and prediction intervals for $E(N_{S})$ and $N_{S}$ using $\sigma_{E(n_{S})}^{2}$ and the variance directly observed from the simulations, respectively.
 
 # Results
-The methods described here for calculating $P({S_{j}})$ and $E(N_{S})$ performed well. 95% confidence intervals for simulated loci overlapped the simulated mean in
+The methods described here for calculating $P({S_{j}})$ and $E(N_{S})$ performed well.
+Individual $P({S_{jq}})$ values for each locus were within the confidence intervals derived from the simulated values for 97 and 96% of loci from the $n = 100$ and $n = 1000$ populations, respectively (Figure 1).
+Note that variation along the generally correlated minor allele frequency/$P({S_{jq}})$ axes is due to variations in genotype frequencies in the simulated data for a given minor allele frequency.
+$P({S_{jq}})$ values estimated using equation 4 account for this adequately.
+
+$E(N_{S})$ values estimated with equation 5 were similarly accurate and were within the 95% confidence intervals produced using the simulated $N_{S}$ for both population sizes (Figure 2).
+Likewise, the 95% prediction intervals calculated using $\sigma_{E(n_{S})}^{2}$ via equation 7 contained 96.1 and 95.3% of the simulated $N_{S}$ values for $n = 100$ and $n = 1000$, respectively (Figure 2).
+
+# Discussion
+
+# References

### 2023-12-20 scripts/seg_sites_rarefaction.Rmd
Introduction draft added

@@
 ## define parameters
-seed <- 231232
+seed <- 231233
 allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
@@
 populate <- function(g, missing, minors){
+  if(length(missing) > 0){
     g[missing] <- NA
     g[-missing][minors] <- 1
+  }
+  else{
+    g[minors] <- 1
+  }
   return(g)
@@
 
-# taken from the binom package, asymptotic method
-confint_pb <- function(p, n, alpha){
-  alpha <- rep(alpha, length = length(p))
-  alpha2 <- 0.5 * alpha
-  z <- qnorm(1 - alpha2)
-  z2 <- z * z
-  se <- sqrt(p * (1 - p)/n)
-  lcl <- p - z * se
-  ucl <- p + z * se
-
-  lcl[lcl < 0] <- 0
-  ucl[ucl > 1] <- 1
-  return(data.frame(lower = lcl, upper = ucl))
-}
-
 c1 <- counts(g1) # geno counts
@@
 vseg2 <- varseg(pseg2)
-ciseg1 <- confint_pb(pseg1, k, .05)
-ciseg2 <- confint_pb(pseg2, k, .05)
 
@@
 mpseg1 <- binom::binom.confint(rowSums(seg1), rep(mcmcs, nrow(seg1)), 
-                               method = "exact")[,c("mean", "lower", "upper")]
+                               method = "agresti-coull")[,c("mean", "lower", "upper")]
 mpseg2 <- binom::binom.confint(rowSums(seg2), rep(mcmcs, nrow(seg1)),
-                               method = "exact")[,c("mean", "lower", "upper")]
+                               method = "agresti-coull")[,c("mean", "lower", "upper")]
 
@@
 ```{r}
-# individual loci
+colors <- khroma::color("highcontrast")(3)
+ci_col <- as.character(colors[3])
+colors <- as.character(colors[c(2,1)])
+
+#========plot the emperical CIs around every loci vs the predicted value==============
+# plotting the emperical CIs vs the estiamted mean
 pseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
@@
 
-ggplot(seg[seg$source == "Mathematical",],
+f1 <- ggplot(seg[seg$source == "Mathematical",],
              aes(x = af, y = p)) +
   geom_point(aes(color = in_ci)) +
-  geom_errorbar(data = cis, aes(ymax = upper, ymin = lower)) +
+  geom_errorbar(aes(ymax = upper, ymin = lower)) +
   facet_wrap(~pop) +
-  theme_bw()
+  theme_bw() +
+  theme(strip.background = element_blank()) +
+  scale_color_manual(values = colors) +
+  guides(color = guide_legend(title = "In CI")) +
+  ylab(bquote(italic(P*"("*S[jq]*")"))) +
+  xlab("Minor Allele Frequency")
+
+ggsave("Figure1.pdf", f1, width = 14, height = 7)
+
+tapply(seg[seg$source == "Mathematical", "in_ci"], seg$pop[seg$source == "Mathematical"], mean) # percent in interval
+#===============plot all of the observed values vs the estimated value for the sums=============
+# plot the estimated prediction interval vs the observations and the confint from the simultations
+nap_pi <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)*sqrt(1 + (1/n))
+nap_ci <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)/sqrt(n)
 
@@
 
+# CI/PI
+msum_ci1 <- nap_ci(.05, var(msum1), length(msum1))
+msum_ci2 <- nap_ci(.05, var(msum2), length(msum2))
+
+sum_pi1 <- nap_pi(0.05, sum(vseg1), nal)
+sum_pi2 <- nap_pi(0.05, sum(vseg2), nal)
+
+# add to table
 msum_tab <- data.frame(nseg = c(mean(msum1), mean(msum2)),
-                       lower = c(quantile(msum1, .05), quantile(msum2, .05)),
-                       upper = c(quantile(msum1, .95), quantile(msum2, .95)),
+                       lower = c(mean(msum1) - msum_ci1, mean(msum2) - msum_ci2),
+                       upper = c(mean(msum1) + msum_ci1, mean(msum2) + msum_ci2),
                        pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)))
@@
 colnames(sum_tab)[1] <- "nseg"
+sum_tab <- merge(sum_tab, msum_tab[,c("pop", "lower", "upper")], by = "pop")
+sum_tab$in_ci <- sum_tab$nseg <= sum_tab$upper & sum_tab$nseg >= sum_tab$lower
 
+# add pi
 all_mcmc_sums <- rbind(cbind.data.frame(nseg = msum1, pop = paste0("n = ", n1/2)),
                    cbind.data.frame(nseg = msum2, pop = paste0("n = ", n2/2)))
+pi_tab <- data.frame(pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)),
+                     lower = c(sum_tab$nseg[sum_tab$pop == paste0("n = ", n1/2)] - sum_pi1,
+                               sum_tab$nseg[sum_tab$pop == paste0("n = ", n2/2)] - sum_pi2),
+                     upper = c(sum_tab$nseg[sum_tab$pop == paste0("n = ", n1/2)] + sum_pi1,
+                               sum_tab$nseg[sum_tab$pop == paste0("n = ", n2/2)] + sum_pi2))
+all_mcmc_sums <- merge(all_mcmc_sums, pi_tab, by = "pop")
+all_mcmc_sums$in_pi <- all_mcmc_sums$nseg <=  all_mcmc_sums$upper & all_mcmc_sums$nseg >=  all_mcmc_sums$lower
 
-ggplot(sum_tab, aes(x = pop, y = nseg)) + 
-  geom_point(color = "red") +
-  geom_errorbar(data = msum_tab, aes(ymax = upper, ymin = lower, x= pop), inherit.aes = FALSE) +
-  geom_jitter(data = all_mcmc_sums, height = 0, alpha = .005) +
+Figure2 <- ggplot(sum_tab, aes(x = pop, y = nseg, color = in_ci)) + 
+  geom_jitter(data = all_mcmc_sums, aes(x = pop, y = nseg, color = in_pi), height = 0, alpha = .2) +
+  geom_point(size = 4, shape = 17) +
+  geom_errorbar(data = sum_tab, aes(ymax = upper, ymin = lower, x = pop), 
+                inherit.aes = FALSE, width = .25, color = ci_col, linewidth = 1) +
+  geom_errorbar(data = pi_tab, aes(ymax = upper, ymin = lower, x = pop), 
+                inherit.aes = FALSE, width = .5, linetype = "dashed", color = ci_col, linewidth = 1) +
   theme_bw() +
-  scale_y_continuous(limits = c(0, nal))
+  theme(strip.background = element_blank()) +
+  scale_color_manual(values = colors) +
+  guides(color = guide_legend(title = "In CI/PI")) +
+  ylab(bquote(italic(E*"("*N[S]*")"))) +
+  xlab("Population")
 
-tapply(seg[seg$source == "Mathematical",]$in_ci, seg[seg$source == "Mathematical",]$pop, mean, na.rm = TRUE)
+ggsave("Figure2.pdf", Figure2, height = 7, width = 14)
+
+tapply(all_mcmc_sums$in_pi, all_mcmc_sums$pop, mean, na.rm = TRUE)
+
+#==========others==================
 
@@
                                        pop = paste0("n = ", n2/2)))
+```
 
-# estimated PI -- normal approx works well
-pi1 <- qnorm(.975)*sqrt(sum(vseg1))*sqrt(1 + (1/nal))
-pi2 <- qnorm(.975)*sqrt(sum(vseg2))*sqrt(1 + (1/nal))
-pi <- data.frame(pop = paste0("n = ", n1/2), 
-                 lower = sum(pseg1) - pi1,
-                 upper = sum(pseg1) + pi1)
-pi <- rbind(pi, data.frame(pop = paste0("n = ", n2/2), 
-                 lower = sum(pseg2) - pi2,
-                 upper = sum(pseg2) + pi2))
+# Misc
+```{r}
+# show difference in HWE divergence with different combinations of 5 minor alleles in 100 genotypes
+exd <- rbind(c(rep(0, 45), rep(1, 5)),
+             c(rep(0, 46), 2, rep(1, 3)),
+             c(rep(0, 47), rep(2, 2), 1))
+exds <- import.snpR.data(as.data.frame(exd), snp.meta = data.frame(snp = 1:3), mDat = -9)
+exds <- calc_hwe(exds)
+get.snpR.stats(exds, stats = "hwe")
 
-all_mcmc_sums <- as.data.table(all_mcmc_sums)
-all_mcmc_sums[pop == pi$pop[1], 
-              in_pi := nseg <= pi$upper[1] & 
-                nseg >= pi$lower[1]]
-all_mcmc_sums[pop == pi$pop[2], 
-              in_pi := nseg <= pi$upper[2] & 
-                nseg >= pi$lower[2]]
-
-ggplot(all_mcmc_sums, aes(x = pop, y = nseg, color = in_pi)) +
-  geom_jitter(height = 0, alpha = .5) +
-  geom_errorbar(data = pi, aes(ymax = upper, ymin = lower, x = pop), inherit.aes = FALSE)
-
-tapply(all_mcmc_sums$in_pi, all_mcmc_sums$pop, mean)
+exc <- counts(exd)
+rarefact(exc, 10)
 ```

### 2023-12-20 manuscript/ms.Rmd
Now with short, focused discussion

@@
 The average allele count per locus (which is usually corrected to allelic richness [@kalinowskiCountingAllelesRarefaction2004]) has historically been one of the most prevalent allele frequency independent measures of genetic diversity.
-While this is a particularly useful measure in microsatellite and other datasets utilizing heavily poly-allelic markers, it is less so in single-nucleotide polymorphism (SNP) datasets.
+While this is a particularly useful measure in microsatellite and other datasets utilizing heavily polyallelic markers, it is less so in single-nucleotide polymorphism (SNP) datasets.
 SNP datasets are usually primarily biallelic, and thus the count of the number of alleles per locus in such datasets is less variable across populations and therefore less interesting.
@@
 These estimators are currently implemented and available in the `snpR` R package via the function `calc_seg_sites` [@hemstromSnpRUserFriendly2023], where they can be automatically run for all populations either with automatic, per-locus optimized control of target sample size or globally via manual control.
-
-<!-- Outline: -->
-
-<!-- * What is genetic diveristy? -->
-<!--   - Utility -->
-<!--   - Estimators -->
-<!--     - heterozygosity -->
-<!--     - pi -->
-<!--     - private alleles -->
-<!--     - allele counts per locus -->
-
-<!-- * Problems with ac and private alleles per locus -->
-<!--   - All subject to bias due to differences in effective sample size -->
-<!--     - Missing data, unequal sample sizes -->
-
-<!-- * Rarefaction -->
-<!--   - Sampling vs mathematical solutions -->
-<!--   - solutions for private alleles and allele counts (richness) -->
-<!--   - description -->
-<!--   - how this solves problems -->
-
-<!-- * Nseg -->
-<!--   - Newer metric for modern sequencing tech -->
-<!--   - Useful alternative, lets you capture different parts of variance -->
-<!--   - biases -->
-<!--   - rarefaction as solution -->
 
@@
 
-To remedy these problems, I propose the following estimator of $P({S_{j}})$:
+To remedy these problems, we propose the following estimator of $P({S_{j}})$:
 $$P({S_{j}}) = 1 - \sum_{k = 1}^{h}\frac{\binom{n_{kj}}{\gamma}}{\binom{n_{j}}{\gamma}}$$
@@
 ## Emperical Validation
-To validate equations 4 and 5, I simulated genotypic data for two populations with sizes 100 and 1000, each with 100 bi-allelic loci with minor allele frequencies spaced equally between 0.01 and 0.1.
-I added missing data to each population assigning each locus a missing data rate $R_{mq}$ from a uniform distribution such that $R_{mq} \sim U(0,.3)$, ensuring that overall allele frequencies in each population were maintained.
-I then used the methods described above to estimate $P({S_{jq}})$ and $E(N_{S})$ and their variances given $\gamma = 30$.
-For comparison, I also conducted 10,000 random draws of size $\gamma$ from each loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws.
-I likewise calculated the variance of $N_{S}$ directly across all sets of random draws.
+To validate equations 4 and 5, we simulated genotypic data for two populations with sizes 100 and 1000, each with 100 bi-allelic loci with minor allele frequencies spaced equally between 0.01 and 0.1.
+We added missing data to each population assigning each locus a missing data rate $R_{mq}$ from a uniform distribution such that $R_{mq} \sim U(0,.3)$, ensuring that overall allele frequencies in each population were maintained.
+We then used the methods described above to estimate $P({S_{jq}})$ and $E(N_{S})$ and their variances given $\gamma = 30$.
+For comparison, we also conducted 10,000 random draws of size $\gamma$ from each loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws.
+We likewise calculated the variance of $N_{S}$ directly across all sets of random draws.
 
-To compare my calculated estimates to the empirical simulations, I used the implementation of the Agresti-Coull [@agrestiApproximateBetterExact1998] method from the R package `binom` [@dorai-rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each empirical $P({S_{jq}})$.
-I also used normal approximation to calculate 95% confidence and prediction intervals for $E(N_{S})$ and $N_{S}$ using $\sigma_{E(n_{S})}^{2}$ and the variance directly observed from the simulations, respectively.
+To compare our calculated estimates to the empirical simulations, we used the implementation of the Agresti-Coull [@agrestiApproximateBetterExact1998] method from the R package `binom` [@dorai-rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each empirical $P({S_{jq}})$.
+We also used normal approximation to calculate 95% confidence and prediction intervals for $E(N_{S})$ and $N_{S}$ using $\sigma_{E(n_{S})}^{2}$ and the variance directly observed from the simulations, respectively.
 
@@
 # Discussion
+The number of segregating sites can provide an excellent counterpoint to measures of genetic diversity that are based on allele frequencies (such as observed and expected heterozygosity).
+For example, populations that have experienced a recent population expansion will often carry an excess of low frequency variants [@gattepailleInferringPopulationSize2013] caused by the recent increase in the overall rate at which mutations are produced in the population and the relative lack of time for any such new variants to drift to higher frequencies.
+The average expected heterozygosity across segregating sites may actually be lower (or at least lower than expected) in such cases than populations which have been demographically static, but they will carry far more segregating sites.
+A recent study in yellow perch provides an excellent example of such: several recently expanded populations show a relatively slight difference in heterozygosity in comparison to other, more demographically static populations but are segregating at far more loci [@schraidtDispersiveCurrentsExplain2023].
+In cases like this, it is particularly important that the number of segregating sites is being properly calculated across populations, since failing to correct for differences in sample size or data missingness could mask biologically interesting signals of demographic history and obscure a critical facet of overall genomic diversity.
+The method we present here should therefore be a useful for future studies of genetic diversity.
 

### 2024-01-23 manuscript/ms.Rmd
First completed figure/text draft

@@
 
+Population geneticists use a wide range of approaches to measure population genetic diversity.
+While measures which incorporate allele frequencies (such as observed and expected heterozygosity, nucleotide diversity, Tajima's D, and so on) are well developed for single nucleotide polymorphism (SNP) data, there are only a few allele frequency independent estimators which are broadly in use.
+The number of segregating sites (or SNPs) is one of the more prevalent of these; however, unlike allelic richness or private allele frequencies, segregating site counts are rarely adjusted to correct for unequal sample sizes or missing data rates between populations.
+We here propose a corrected estimator for the number of segregating sites expected in a population following rarefaction, which allows for an unbiased approach for comparing the number of segregating sites between multiple populations.
+
 # Introduction
@@
 
-Estimators of population genetic diversity are themselves extremely diverse, but can essentially be split into two categories depending on if they are allele frequency dependent or independent.
-Allele frequency dependent statistics, such as expected heterozygosity ($H_{e}$), the F-statistics ($F_{IS}$, $F_{ST}$, and $F_{IT}$)[@Weir1984], nucleotide diversity ($\pi$), and Tajima's $\theta$ [@Tajima1989] all rely principally on the frequency of the alleles across any surveyed loci; whereas allele frequency independent statistics, such as Watterson's $\theta$ [@wattersonNumberSegregatingSites1975], counts of alleles per loci, the number of observed segregating sites, and (to some degree) private allele counts do not.
+Estimators of population genetic diversity are themselves extremely diverse, but essentially fall into two categories: those that are depend on allele frequencies and those that do not.
+Allele frequency dependent statistics, such as expected heterozygosity ($H_{e}$), the F-statistics ($F_{IS}$, $F_{ST}$, and $F_{IT}$)[@Weir1984], nucleotide diversity ($\pi$), and Tajima's $\theta$ [@Tajima1989] all rely principally on the frequency of the alleles across any surveyed loci; whereas allele frequency independent statistics, such as Watterson's $\theta$ [@wattersonNumberSegregatingSites1975], counts of alleles per loci, the number of observed segregating sites, and private allele counts do not.
 This is not a trivial difference:  evolutionary forces do not act equally on the number of loci/alleles in a population and the frequency of those alleles.
-Tajima's D, for example, is a powerful and broadly used statistic fundamentally based on the difference between Tajima's $\theta$ and Watterson's $\theta$ can be used to both detect selection and population demographic changes [@Tajima1989] due to the way that those forces act on the balance of allele frequencies within populations.
+Tajima's D, for example, is a powerful and broadly used statistic fundamentally based on the difference between Tajima's $\theta$ and Watterson's $\theta$ that can be used to both detect selection and population demographic changes [@Tajima1989] due to the way that those forces act on the balance of allele frequencies within populations.
 
@@
 While this is a particularly useful measure in microsatellite and other datasets utilizing heavily polyallelic markers, it is less so in single-nucleotide polymorphism (SNP) datasets.
-SNP datasets are usually primarily biallelic, and thus the count of the number of alleles per locus in such datasets is less variable across populations and therefore less interesting.
-Counts of the number of segregating sites per population are a useful alternative in SNP datasets given the large number of possible segregating sites that these datasets often query [@?CITE].
+SNP datasets are usually biallelic by design, and thus allele counts per locus vary little populations and are therefore less informative.
+The number of segregating sites per population is a useful alternative in SNP datasets given the large number of possible segregating sites that these datasets often query [@?CITE].
 
-However, raw counts of the number of segregating sites per population are problematic whenever sample sizes are not equal across the populations under comparison at all loci.
+However, segregating sites counts per population are problematic whenever sample sizes are not equal across the populations under comparison at all loci.
 This can occur either directly due to unequal numbers of individuals or  indirectly due to unqual proportions of missing data across populations.
@@
 # Discussion
-The number of segregating sites can provide an excellent counterpoint to measures of genetic diversity that are based on allele frequencies (such as observed and expected heterozygosity).
+We present here an method to correct for the probability that a given loci would be observed as segregating following rarefaction to a given sample size.
+This provides for the straightforward estimation of the total number of segregating sites which would be observed in a population sample at any reduced size, and therefore provides a way to standardize that metric across samples from different populations.
+We also show that this approach is unbiased by allele frequency or missing data variation across individual samples or loci.
+
+Our estimator should useful given that the number of segregating sites can provide an excellent counterpoint to measures of genetic diversity that are based on allele frequencies (such as observed and expected heterozygosity).
 For example, populations that have experienced a recent population expansion will often carry an excess of low frequency variants [@gattepailleInferringPopulationSize2013] caused by the recent increase in the overall rate at which mutations are produced in the population and the relative lack of time for any such new variants to drift to higher frequencies.
@@
 A recent study in yellow perch provides an excellent example of such: several recently expanded populations show a relatively slight difference in heterozygosity in comparison to other, more demographically static populations but are segregating at far more loci [@schraidtDispersiveCurrentsExplain2023].
+
 In cases like this, it is particularly important that the number of segregating sites is being properly calculated across populations, since failing to correct for differences in sample size or data missingness could mask biologically interesting signals of demographic history and obscure a critical facet of overall genomic diversity.
-The method we present here should therefore be a useful for future studies of genetic diversity.
+The $E(N_{S})$ method we present here should therefore be a useful for future studies of genetic diversity, and is currently available for use via the function "calc_seg_sites()" in the R package "snpR" [@hemstromSnpRUserFriendly2023].
 
 # References
+
+# Figure Captions:
+
+Figure 1: The expected probabilities $P({S_{j,a}})$ of observing a segregating site at a locus for loci with different minor allele frequencies.
+Probabilities are derived from Equation 4 for each locus *a* for population sizes of 100 (left) and 1000 (right), each rarefacted to $n = 30$.
+Each locus has a random, independent percentage of missing data which varies between 0 and 30.
+Points are colored depending on if they fell within a 95% confidence level based on 10,000 random re-samplings of each population to 30 individuals.
+
+Figure 2: The expected total number of segregating sites $E(N_{S})$ from populations of either 100 or 1,000 sampled to 30 individuals as calculated via Equation 5 vs that observed from 10,000 random re-samplings to 30 individuals.
+95% Confidence and prediction intervals for $E(N_{S})$ are shown, calculated according to Equations 8 and 9.
+Observed $N_{S}$ values from each re-sample are colored depending on if they fell within the estimated 95% prediction interval.
+The mean $N_{S}$ across all re-samples is marked with a triangle, colored depending on if it fell within the estimated 95% confidence interval.

### 2024-01-23 seg_sites_rarefaction.R

@@ -1,0 +1,225 @@
+## ----setup, include=FALSE------------------------------------------------------------------------------------------------------
+knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table)
+
+
+
+## ------------------------------------------------------------------------------------------------------------------------------
+# Generate some example data:
+## define parameters
+seed <- 231233
+allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
+n1 <- 2000 # p1 gene copy number
+n2 <- 200 # p2 gene copy number
+k <- 30 # number to rarefact to
+mcmcs <- 1000 # number of mcmc draws
+max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
+
+
+## derived
+set.seed(seed)
+nal <- length(allele_frequencies)
+miss_percent <- floor(runif(nal, 0, max_missing))/100
+
+
+
+## ------------------------------------------------------------------------------------------------------------------------------
+# generate genotypes
+g1 <- matrix(0, nrow = nal, ncol = n1) # init
+g2 <- matrix(0, nrow = nal, ncol = n2)
+
+# add missing data
+ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
+ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
+
+# determine minor allele placements
+ma1 <- ma2 <- vector("list", length(ms1))
+for(i in 1:nal){
+  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
+  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)
+
+}
+
+# populate
+## function to replace minors
+populate <- function(g, missing, minors){
+  if(length(missing) > 0){
+    g[missing] <- NA
+    g[-missing][minors] <- 1
+  }
+  else{
+    g[minors] <- 1
+  }
+  return(g)
+}
+## do for every locus
+for(i in 1:nal){ 
+  g1[i,] <- populate(g1[i,], ms1[[i]], ma1[[i]])
+  g2[i,] <- populate(g2[i,], ms2[[i]], ma2[[i]])
[…]

### 2024-02-13 scripts/seg_sites_rarefaction.Rmd
additional simulations spanning a bigger range of K and mcmc count, new figures in progress

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table)
+knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table);
+```
 
-```
 # Define parameters
 ```{r}
-# Generate some example data:
-## define parameters
+# define parameters
 seed <- 231233
+threads <- 25
 allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
-n1 <- 2000 # p1 gene copy number
-n2 <- 200 # p2 gene copy number
-k <- 30 # number to rarefact to
-mcmcs <- 10000 # number of mcmc draws
+n1 <- 5000 # p1 gene copy number
+n2 <- 500 # p2 gene copy number
+k <- seq(10, 100, by = 10) # numbers to rarefact to
+mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
 max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
@@
 
+## plotting
+colors <- khroma::color("highcontrast")(3)
+ci_col <- as.character(colors[3])
+colors <- as.character(colors[c(2,1)])
+t.sizes <- c(20, 18)
 ```
 
-# Generate data
+# Setup genotypes
 ```{r}
@@
 g2 <- g2[,1:(ncol(g2)/2)] + g2[,((ncol(g2)/2) + 1):ncol(g2)]
+
 ```
 
-# Via rarefaction approach
+# Rarefact mathematically
 ```{r}
@@
 rarefact <- function(counts, k){
+  if(length(k) > 1){
+    warning("length(k) > 1, using first element only.\n")
+    k <- k[1]
+  }
   return(1 - ((choose(counts[,1], k) + choose(counts[,3], k))/choose(rowSums(counts), k)))
@@
 
-pseg1 <- rarefact(c1, k) # rarefact
-pseg2 <- rarefact(c2, k) # rarefact
+# rarefact
+pseg1 <- lapply(k, function(tk) rarefact(c1, tk))
+pseg2 <- lapply(k, function(tk) rarefact(c2, tk))
+names(pseg1) <- names(pseg2) <- paste0("K_", k)
 
-vseg1 <- varseg(pseg1)
-vseg2 <- varseg(pseg2)
+vseg1 <- lapply(pseg1, varseg)
+vseg2 <- lapply(pseg2, varseg)
 
-p1 <- sum(pseg1)/nal
-p2 <- sum(pseg2)/nal
+p1 <- lapply(pseg1, function(p) sum(p)/nal)
+p2 <- lapply(pseg2, function(p) sum(p)/nal)
 ```
 
-# Via mcmc
+# Rarefact via simulations
+Note: this chunck should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunck.
+`q` should be passed as an argument for the iteration, as should `outfile`.
 
-```{r}
+```{r cluster_MCMCs, eval=FALSE}
+set.seed(seed + q) # reset seed for cluster
+
 rarefact_mcmc <- function(g, k){
+  if(length(k) > 1){
+    warning("length(k) > 1, using first element only.\n")
+    k <- k[1]
+  }
+
   miss <- is.na(g)
@@
 
-# draw
-seg1 <- seg2 <- matrix(0, nal, mcmcs)
-for(i in 1:mcmcs){
+# draw for each k/mcmc count combination
+combs <- expand.grid(k, mcmcs)
+colnames(combs) <- c("k", "iters")
+
+# draw for each k/mcmc count combination
+combs <- expand.grid(k, mcmcs)
+colnames(combs) <- c("k", "iters")
+
+tmcmcs <- combs$iters[q]
+tk <- combs$k[q]
+
+seg1 <- seg2 <- matrix(0, nal, tmcmcs)
+for(i in 1:tmcmcs){
   for(j in 1:nal){
-    seg1[j,i] <- rarefact_mcmc(g1[j,], k = k)
-    seg2[j,i] <- rarefact_mcmc(g2[j,], k = k)
+    seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
+    seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
   }
@@
 
-mpseg1 <- binom::binom.confint(rowSums(seg1), rep(mcmcs, nrow(seg1)), 
+mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
                                method = "agresti-coull")[,c("mean", "lower", "upper")]
-mpseg2 <- binom::binom.confint(rowSums(seg2), rep(mcmcs, nrow(seg1)),
+mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
                                method = "agresti-coull")[,c("mean", "lower", "upper")]
@@
 
+res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
+
+saveRDS(res, paste0(outfile, "_", q, ".RDS"))
+
 ```
 
-# results
+# import and condense simulation results
 ```{r}
-colors <- khroma::color("highcontrast")(3)
-ci_col <- as.character(colors[3])
-colors <- as.character(colors[c(2,1)])
+res_files <- list.files("../data/seg_sites/", "\\.RDS", full.names = TRUE)
+res_files <- lapply(res_files, readRDS)
+ks <- unlist(purrr::map(res_files, "k"))
+n_iters <- unlist(purrr::map(res_files, "iters"))
 
+```
+
+# Figure 1: Pseg for four individual runs (k = 10 and 100, N = 100, N = 1000)
+
+```{r Figure1}
+prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2){
   #========plot the emperical CIs around every loci vs the predicted value==============
   # plotting the emperical CIs vs the estiamted mean
-pseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
-              data.frame(pop = paste0("n = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))
+  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
+                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))
 
-oseg <- rbind(data.frame(pop = paste0("n = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies),
-              data.frame(pop = paste0("n = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies))
+  oseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies),
+                data.frame(pop = paste0("N = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies))
 
@@
   seg <- rbind(pseg, oseg)
-cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("n = ", n1/2), af = allele_frequencies),
-             cbind.data.frame(mpseg2, pop = paste0("n = ", n2/2), af = allele_frequencies))
+  cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("N = ", n1/2), af = allele_frequencies),
+               cbind.data.frame(mpseg2, pop = paste0("N = ", n2/2), af = allele_frequencies))
   colnames(cis) <- c("p", "lower", "upper", "pop", "af")
@@
 
-f1 <- ggplot(seg[seg$source == "Mathematical",],
+  return(seg)
+}
+
+K10_data <- which(ks == 10 & n_iters == 10000)
+F1data_K10 <- prep_f1(pseg1 = pseg1$K_10, 
+                      pseg2 = pseg2$K_10, 
+                      mpseg1 = res_files[[K10_data]]$mpseg1,
+                      mpseg2 = res_files[[K10_data]]$mpseg2)
+
+K100_data <- which(ks == 100 & n_iters == 10000)
+F1data_K100 <- prep_f1(pseg1 = pseg1$K_100, 
+                       pseg2 = pseg2$K_100, 
+                       mpseg1 = res_files[[K100_data]]$mpseg1,
+                       mpseg2 = res_files[[K100_data]]$mpseg2)
+
+F1data_K10$K <- "gamma * ' = 10'"
+F1data_K100$K <- "gamma * ' = 100'"
+
+F1data <- rbind(F1data_K10, F1data_K100)
+F1data$pop <- paste0("'", F1data$pop, "'")
+
+f1 <- ggplot(F1data[F1data$source == "Mathematical",],
              aes(x = af, y = p)) +
@@
   geom_errorbar(aes(ymax = upper, ymin = lower)) +
-  facet_wrap(~pop) +
+  facet_grid(K~pop, labeller = label_parsed) +
   theme_bw() +
@@
   ylab(bquote(italic(P*"("*S[jq]*")"))) +
-  xlab("Minor Allele Frequency")
+  xlab("Minor Allele Frequency") +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        legend.title = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2]))
 
-ggsave("Figure1.pdf", f1, width = 14, height = 7)
+ggsave("Figure1.pdf", f1, width = 15, height = 10)
+tapply(F1data[F1data$source == "Mathematical", "in_ci"], F1data[F1data$source == "Mathematical", c("K", "pop")], mean) # percent in interval
+```
 
-tapply(seg[seg$source == "Mathematical", "in_ci"], seg$pop[seg$source == "Mathematical"], mean) # percent in interval
-#===============plot all of the observed values vs the estimated value for the sums=============
-# plot the estimated prediction interval vs the observations and the confint from the simultations
+# Figure 2: Distribution of Nseg vs Eseg for four runs
+
+```{r}
+prep_f2 <- function(seg1, seg2, vseg1, vseg2, F1data){
+  seg <- F1data
+
   nap_pi <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)*sqrt(1 + (1/n))
@@
   sum_tab <- as.data.frame(sum_tab)
-sum_tab$pop <- rownames(sum_tab)
+  sum_tab$pop <- tolower(rownames(sum_tab))
   colnames(sum_tab)[1] <- "nseg"
@@
 
+  return(list(sum_tab = sum_tab, all_mcmc_sums = all_mcmc_sums, pi_tab = pi_tab))
+}
+
+K10_data <- which(ks == 10 & n_iters == 10000)
+F2data_K10 <- prep_f2(seg1 = res_files[[K10_data]]$seg1, 
+                      seg2 = res_files[[K10_data]]$seg2, 
+                      vseg1 = vseg1$K_10,
+                      vseg2 = vseg2$K_10,
+                      F1data = F1data_K10)
+
+K100_data <- which(ks == 100 & n_iters == 10000)
+F2data_K100 <- prep_f2(seg1 = res_files[[K100_data]]$seg1, 
+                       seg2 = res_files[[K100_data]]$seg2, 
+                       vseg1 = vseg1$K_100,
+                       vseg2 = vseg2$K_100,
+                       F1data = F1data_K100)
+
+F2data_K10$sum_tab$K <- "gamma * ' = 10'"
+F2data_K100$sum_tab$K <- "gamma * ' = 100'"
+F2data_K10$all_mcmc_sums$K <- "gamma * ' = 10'"
+F2data_K100$all_mcmc_sums$K <- "gamma * ' = 100'"
+F2data_K10$pi_tab$K <- "gamma * ' = 10'"
+F2data_K100$pi_tab$K <- "gamma * ' = 100'"
+
+F2data_sums <- rbind(F2data_K10$sum_tab, F2data_K100$sum_tab)
+F2data_sums$pop <- gsub("n", "N", F2data_sums$pop)
+F2data_sums$pop <- paste0("'", F2data_sums$pop, "'")
+
+F2data_mcmcs <- rbind(F2data_K10$all_mcmc_sums, F2data_K100$all_mcmc_sums)
+F2data_mcmcs$pop <- gsub("n", "N", F2data_mcmcs$pop)
+F2data_mcmcs$pop <- paste0("'", F2data_mcmcs$pop, "'")
+
+F2data_pi <- rbind(F2data_K10$pi_tab, F2data_K100$pi_tab)
+F2data_pi$pop <- gsub("n", "N", F2data_pi$pop)
+F2data_pi$pop <- paste0("'", F2data_pi$pop, "'")
+
+f2 <- ggplot(F2data_mcmcs, aes(x = nseg)) +
+  geom_density() +
+  geom_vline(data = F2data_sums, aes(xintercept = nseg), color = colors[1]) +
+  geom_vline(data = F2data_pi, aes(xintercept = lower), linetype = "dashed", color = colors[2]) +
+  geom_vline(data = F2data_pi, aes(xintercept = upper), linetype = "dashed", color = colors[2]) +
+  facet_grid(K~pop, labeller = label_parsed) +
+  theme_bw() +
+  scale_y_continuous(expand = c(0,0)) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        strip.background = element_blank()) +
+  xlab(bquote(italic(E*"("*N[S]*")"))) +
+  ylab("Density")
+
 Figure2 <- ggplot(sum_tab, aes(x = pop, y = nseg, color = in_ci)) + 
@@
 
-ggsave("Figure2.pdf", Figure2, height = 7, width = 14)
+ggsave("Figure2.pdf", f2, height = 10, width = 15)
 
 tapply(all_mcmc_sums$in_pi, all_mcmc_sums$pop, mean, na.rm = TRUE)
+```
+
+```{r}
+
+#===============plot all of the observed values vs the estimated value for the sums=============
+# plot the estimated prediction interval vs the observations and the confint from the simultations
 
@@
 
-# Misc
 ```{r}

### 2024-02-13 data/seg_sites/seg_sites_rarefaction.R

@@ -1,0 +1,140 @@
+args <- commandArgs(TRUE)
+q <- as.numeric(args[1])
+outfile <- as.character(args[2])
+
+library(snpR); library(ggplot2); library(data.table);
+
+## ------------------------------------------------------------------------------------------------------------------------------
+# Generate some example data:
+## define parameters
+seed <- 231233
+threads <- 25
+allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
+n1 <- 5000 # p1 gene copy number
+n2 <- 500 # p2 gene copy number
+k <- seq(10, 100, by = 10) # numbers to rarefact to
+mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
+max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
+
+
+## derived
+set.seed(seed)
+nal <- length(allele_frequencies)
+miss_percent <- floor(runif(nal, 0, max_missing))/100
+
+
+## ------------------------------------------------------------------------------------------------------------------------------
+# generate genotypes
+g1 <- matrix(0, nrow = nal, ncol = n1) # init
+g2 <- matrix(0, nrow = nal, ncol = n2)
+
+# add missing data
+ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
+ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
+
+# determine minor allele placements
+ma1 <- ma2 <- vector("list", length(ms1))
+for(i in 1:nal){
+  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
+  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)
+
+}
+
+# populate
+## function to replace minors
+populate <- function(g, missing, minors){
+  if(length(missing) > 0){
+    g[missing] <- NA
+    g[-missing][minors] <- 1
+  }
+  else{
+    g[minors] <- 1
+  }
+  return(g)
+}
+## do for every locus
+for(i in 1:nal){ 
[…]

### 2024-02-13 data/seg_sites/seg_sites_rarefaction.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+#SBATCH --job-name=seg_sites
+#SBATCH -A standby
+#SBATCH --mem=20G
+#SBATCH --array=1-100
+#SBATCH -t 4:00:00
+
+module purge
+module load r/4.2
+
+outfile=mcmc_res
+
+Rscript seg_sites_rarefaction.R $SLURM_ARRAY_TASK_ID $outfile

### 2024-02-13 scripts/seg_sites_rarefaction.R

@@ -1,17 +1,17 @@
 ## ----setup, include=FALSE------------------------------------------------------------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table)
-
+knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table); library(foreach); library(doParallel)
 
 
 ## ------------------------------------------------------------------------------------------------------------------------------
 # Generate some example data:
 ## define parameters
 seed <- 231233
+threads <- 25
 allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
-n1 <- 2000 # p1 gene copy number
-n2 <- 200 # p2 gene copy number
-k <- 30 # number to rarefact to
-mcmcs <- 1000 # number of mcmc draws
+n1 <- 5000 # p1 gene copy number
+n2 <- 500 # p2 gene copy number
+k <- seq(10, 100, by = 10) # numbers to rarefact to
+mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
 max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
 
 
@@ -19,7 +19,6 @@ max_missing <- 30 # maximum percentage of missing data in a locus, from a unifor
 set.seed(seed)
 nal <- length(allele_frequencies)
 miss_percent <- floor(runif(nal, 0, max_missing))/100
-
 
 
 ## ------------------------------------------------------------------------------------------------------------------------------
@@ -62,6 +61,7 @@ g1 <- g1[,1:(ncol(g1)/2)] + g1[,((ncol(g1)/2) + 1):ncol(g1)]
 g2 <- g2[,1:(ncol(g2)/2)] + g2[,((ncol(g2)/2) + 1):ncol(g2)]
 
 
+
 ## ------------------------------------------------------------------------------------------------------------------------------
 counts <- function(g){
   return(data.frame(majhom = rowSums(g == 0, na.rm = TRUE), 
@@ -69,6 +69,10 @@ counts <- function(g){
[…]

### 2024-02-15 scripts/seg_sites_rarefaction.Rmd
work on figure revisions

@@
 knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table);
+library(cowplot)
 ```
@@
 # define parameters
-seed <- 231233
+seed <- 231266
 threads <- 25
@@
 
-# draw for each k/mcmc count combination
-combs <- expand.grid(k, mcmcs)
-colnames(combs) <- c("k", "iters")
-
 tmcmcs <- combs$iters[q]
@@
 ```{r}
-res_files <- list.files("../data/seg_sites/", "\\.RDS", full.names = TRUE)
+res_files <- list.files("../data/", "\\.RDS", full.names = TRUE)
 res_files <- lapply(res_files, readRDS)
@@
 ```{r Figure1}
-prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2){
+prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
   #========plot the emperical CIs around every loci vs the predicted value==============
@@
                       mpseg1 = res_files[[K10_data]]$mpseg1,
-                      mpseg2 = res_files[[K10_data]]$mpseg2)
+                      mpseg2 = res_files[[K10_data]]$mpseg2,
+                      n1 = n1,
+                      n2 = n2)
 
@@
                        mpseg1 = res_files[[K100_data]]$mpseg1,
-                       mpseg2 = res_files[[K100_data]]$mpseg2)
+                       mpseg2 = res_files[[K100_data]]$mpseg2,
+                       n1 = n1,
+                       n2 = n2)
 
@@
 ```{r}
-prep_f2 <- function(seg1, seg2, vseg1, vseg2, F1data){
+prep_f2 <- function(seg1, seg2, vseg1, vseg2, F1data, n1, n2){
   seg <- F1data
 
+  # normal approximation CI and PI equation given var and N
   nap_pi <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)*sqrt(1 + (1/n))
@@
                       vseg2 = vseg2$K_10,
-                      F1data = F1data_K10)
+                      F1data = F1data_K10,
+                      n1 = n1,
+                      n2 = n2)
 
@@
                        vseg2 = vseg2$K_100,
-                       F1data = F1data_K100)
+                       F1data = F1data_K100,
+                       n1 = n1,
+                       n2 = n2)
 
@@
         strip.background = element_blank()) +
-  xlab(bquote(italic(E*"("*N[S]*")"))) +
+  xlab(bquote(italic(N[S]))) +
   ylab("Density")
-
-Figure2 <- ggplot(sum_tab, aes(x = pop, y = nseg, color = in_ci)) + 
-  geom_jitter(data = all_mcmc_sums, aes(x = pop, y = nseg, color = in_pi), height = 0, alpha = .2) +
-  geom_point(size = 4, shape = 17) +
-  geom_errorbar(data = sum_tab, aes(ymax = upper, ymin = lower, x = pop), 
-                inherit.aes = FALSE, width = .25, color = ci_col, linewidth = 1) +
-  geom_errorbar(data = pi_tab, aes(ymax = upper, ymin = lower, x = pop), 
-                inherit.aes = FALSE, width = .5, linetype = "dashed", color = ci_col, linewidth = 1) +
-  theme_bw() +
-  theme(strip.background = element_blank()) +
-  scale_color_manual(values = colors) +
-  guides(color = guide_legend(title = "In CI/PI")) +
-  ylab(bquote(italic(E*"("*N[S]*")"))) +
-  xlab("Population")
 
@@
 
-tapply(all_mcmc_sums$in_pi, all_mcmc_sums$pop, mean, na.rm = TRUE)
+```
+
+# Figure 3: var vs difference between estimate and mean of simulations for mcmc and k
+
+```{r Figure3}
+# for all of the mpseg files, figure out if the percentage of points in the CI changes
+## do for both each SNP and for Ns
+ps_list <- ns_list <- vector("list", length(res_files))
+for(i in 1:length(res_files)){
+  tk <- res_files[[i]]$k
+  titers <- res_files[[i]]$iters
+
+  ps_prep <- prep_f1(pseg1 = pseg1[[paste0("K_", tk)]],
+                     pseg2 = pseg2[[paste0("K_", tk)]], 
+                     mpseg1 = res_files[[i]]$mpseg1,
+                     mpseg2 = res_files[[i]]$mpseg2,
+                     n1 = n1,
+                     n2 = n2)
+
+  ns_prep <- prep_f2(seg1 = res_files[[i]]$seg1, 
+                     seg2 = res_files[[i]]$seg2, 
+                     vseg1 = vseg1[[paste0("K_", tk)]],
+                     vseg2 = vseg2[[paste0("K_", tk)]],
+                     F1data = ps_prep,
+                     n1 = n1,
+                     n2 = n2)
+
+  ps_prep$K <- tk
+  ps_prep$iters <- titers
+  ps_prep$sim_num <- i
+
+  ns_prep$all_mcmc_sums$K <- tk
+  ns_prep$all_mcmc_sums$iters <- titers
+  ns_prep$all_mcmc_sums$i <- i
+  ns_prep$all_mcmc_sums$ENs <- ns_prep$sum_tab$nseg[match(ns_prep$all_mcmc_sums$pop,
+                                                          ns_prep$sum_tab$pop)]
+
+  ps_list[[i]] <- ps_prep
+  ns_list[[i]] <- ns_prep$all_mcmc_sums
+}
+
+ps_all <- rbindlist(ps_list)
+
+ps_all_m <- ps_all[source == "Mathematical",]
+ps_all_s <- ps_all[source == "Simulation",]
+ps_all_merge <- merge(ps_all_m[,-c("source")], 
+                      ps_all_s[,-c("source")], 
+                      by = c("af", "pop", "snp", "sim_num", "lower", "upper", "K", "iters"), 
+                      suffixes = c(".mathematical", ".simulated"))
+ps_all_merge[,diff := p.mathematical - p.simulated]
+
+# Figure S1: gamma vs error rate
+ps_all_aggregate_for_K <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(K, iters, pop)]
+FS1 <- ggplot(ps_all_aggregate_for_K, aes(x = K, y = in_ci.mathematical*100)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_grid(iters~pop) +
+  theme_bw() +
+  # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
+  # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        strip.background = element_blank(),
+        legend.title = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2])) +
+  xlab(parse(text = "gamma")) +
+  ylab("% estimates in CI")
+
+# Figure S2: maf vs error rate
+ps_all_aggregate_for_af <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(af, iters, pop)]
+FS2 <- ggplot(ps_all_aggregate_for_af, aes(x = af, y = in_ci.mathematical*100)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_grid(iters~pop) +
+  theme_bw() +
+  # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
+  # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        strip.background = element_blank(),
+        legend.title = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2])) +
+  xlab("Minor Allele Frequency") +
+  ylab("% estimates in CI")
+
+f3a <- ggplot(ps_all_merge, aes(x = K, y = diff, color = in_ci.mathematical)) +
+  geom_point() +
+  theme_bw() +
+  facet_grid(~pop) +
+  khroma::scale_color_highcontrast() +
+  scale_x_continuous(breaks = k) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        strip.background = element_blank(),
+        legend.title = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2])) +
+  xlab(parse(text = "gamma")) +
+  ylab(bquote(italic(P*"("*S[jq]*")"[E]) - italic(P*"("*S[jq]*")"[O]))) +
+  guides(color = guide_legend(title = "In 95% CI"))
+
+f3b <- ggplot(ps_all_merge, aes(x = af, y = diff, color = in_ci.mathematical)) +
+  geom_point() +
+  theme_bw() +
+  facet_grid(~pop) +
+  khroma::scale_color_highcontrast() +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
+        strip.background = element_blank(),
+        legend.title = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2])) +
+  xlab("Minor Allele Frequency") +
+  ylab(bquote(italic(P*"("*S[jq]*")"[E]) - italic(P*"("*S[jq]*")"[O])))
+
+f3c <- ggplot(ps_all_merge, aes(x = iters, y = diff, color = in_ci.mathematical)) +
+  geom_point() +
+  theme_bw() +
+  facet_grid(~pop) +
+  khroma::scale_color_highcontrast() +
+  scale_x_continuous(breaks = mcmcs) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        axis.text.x = element_text(size = t.sizes[2], 
+                                   angle = 90, vjust = .5),
+        strip.background = element_blank(),
+        legend.title = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2])) +
+  xlab("Number of Simulated Draws") +
+  ylab(bquote(italic(P*"("*S[jq]*")"[E]) - italic(P*"("*S[jq]*")"[O])))
+
+f3l <- ggpubr::get_legend(f3a)
+f3 <- plot_grid(f3a + guides(color = "none"), 
+          f3b + guides(color = "none"), 
+          f3c + guides(color = "none"),
+          ncol = 1, 
+          align = "bl", 
+          axis = "bl")
+
+f3 <- plot_grid(f3, f3l, nrow = 1, rel_widths = c(1, .2))
+
+save_plot("Figure3.pdf", f3, base_height = 11, base_width = 15)
+```
+# Figure 4 Nseg vs parameters
+
+```{r Figure4}
+ns_all <- rbindlist(ns_list)
+
+ns_all$diff <- ns_all$ENs - ns_all$nseg
+
+f4a <- ggplot(ns_all[iters == 10000,], aes(x = K, y = nseg)) +
+  geom_errorbar(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper")]),
+            aes(x = K, ymin = lower, ymax = upper), color = "#DDAA33", size = 1, 
+            position = position_nudge(x = 3.5),
+            inherit.aes = FALSE, width = 2) +
+  stat_ydensity(aes(group = K), draw_quantiles = c(.025, .975), scale = "width", bw = .4,
+                fill = NA, width = 6) +
+  facet_wrap(~pop) +
+  theme_bw() + 
+  scale_x_continuous(breaks = k) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        legend.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        strip.background = element_blank(),
+        legend.text = element_text(size = t.sizes[2])) +
+  ylab(bquote(italic(E*"("*N[S]*")") - italic(N[S,observed]))) +
+  xlab(bquote(gamma))
+
+f4b <- ggplot(ns_all[K == 50,], aes(x = iters, y = nseg)) +
+  geom_errorbar(data = unique(ns_all[K == 50,c("pop", "K", "iters", "i", "lower", "upper")]),
+            aes(x = iters, ymin = lower, ymax = upper), color = "#DDAA33", size = 1, 
+            position = position_nudge(x = 350),
+            inherit.aes = FALSE, width = 200) +
+  stat_ydensity(aes(group = iters), draw_quantiles = c(.025, .975), scale = "width", bw = .4,
+                fill = NA, width = 600) +
+  facet_wrap(~pop) +
+  theme_bw() + 
+  scale_x_continuous(breaks = mcmcs) +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        legend.title = element_text(size = t.sizes[1]),
+        strip.text = element_text(size = t.sizes[1]),
+        strip.background = element_blank(),
+        legend.text = element_text(size = t.sizes[2])) +
+  ylab(bquote(italic(E*"("*N[S]*")") - italic(N[S,observed]))) +
+  xlab("Number of Simulated Draws")
+
 ```

### 2024-02-15 data/seg_sites/seg_sites_rarefaction.R

@@ -1,140 +1,0 @@
-args <- commandArgs(TRUE)
-q <- as.numeric(args[1])
-outfile <- as.character(args[2])
-
-library(snpR); library(ggplot2); library(data.table);
-
-## ------------------------------------------------------------------------------------------------------------------------------
-# Generate some example data:
-## define parameters
-seed <- 231233
-threads <- 25
-allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
-n1 <- 5000 # p1 gene copy number
-n2 <- 500 # p2 gene copy number
-k <- seq(10, 100, by = 10) # numbers to rarefact to
-mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
-max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
-
-
-## derived
-set.seed(seed)
-nal <- length(allele_frequencies)
-miss_percent <- floor(runif(nal, 0, max_missing))/100
-
-
-## ------------------------------------------------------------------------------------------------------------------------------
-# generate genotypes
-g1 <- matrix(0, nrow = nal, ncol = n1) # init
-g2 <- matrix(0, nrow = nal, ncol = n2)
-
-# add missing data
-ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
-ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
-
-# determine minor allele placements
-ma1 <- ma2 <- vector("list", length(ms1))
-for(i in 1:nal){
-  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
-  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)
-
-}
-
-# populate
-## function to replace minors
-populate <- function(g, missing, minors){
-  if(length(missing) > 0){
-    g[missing] <- NA
-    g[-missing][minors] <- 1
-  }
-  else{
-    g[minors] <- 1
-  }
-  return(g)
-}
-## do for every locus
-for(i in 1:nal){ 
[…]

### 2024-02-15 data/seg_sites_rarefaction.R

@@ -1,0 +1,140 @@
+args <- commandArgs(TRUE)
+q <- as.numeric(args[1])
+outfile <- as.character(args[2])
+
+library(snpR); library(ggplot2); library(data.table);
+
+## ------------------------------------------------------------------------------------------------------------------------------
+# Generate some example data:
+## define parameters
+seed <- 231266
+threads <- 25
+allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
+n1 <- 5000 # p1 gene copy number
+n2 <- 500 # p2 gene copy number
+k <- seq(10, 100, by = 10) # numbers to rarefact to
+mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
+max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
+
+
+## derived
+set.seed(seed)
+nal <- length(allele_frequencies)
+miss_percent <- floor(runif(nal, 0, max_missing))/100
+
+
+## ------------------------------------------------------------------------------------------------------------------------------
+# generate genotypes
+g1 <- matrix(0, nrow = nal, ncol = n1) # init
+g2 <- matrix(0, nrow = nal, ncol = n2)
+
+# add missing data
+ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
+ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
+
+# determine minor allele placements
+ma1 <- ma2 <- vector("list", length(ms1))
+for(i in 1:nal){
+  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
+  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)
+
+}
+
+# populate
+## function to replace minors
+populate <- function(g, missing, minors){
+  if(length(missing) > 0){
+    g[missing] <- NA
+    g[-missing][minors] <- 1
+  }
+  else{
+    g[minors] <- 1
+  }
+  return(g)
+}
+## do for every locus
+for(i in 1:nal){ 
[…]

### 2024-04-09 manuscript/ms.Rmd
changes prior to moving to doc for final

@@
 ## Emperical Validation
-To validate equations 4 and 5, we simulated genotypic data for two populations with sizes 100 and 1000, each with 100 bi-allelic loci with minor allele frequencies spaced equally between 0.01 and 0.1.
+To validate equations 4 and 5, we simulated genotypic data for two populations with sizes 250 and 2,500, each with 100 bi-allelic loci with minor allele frequencies spaced equally between 0.01 and 0.1.
 We added missing data to each population assigning each locus a missing data rate $R_{mq}$ from a uniform distribution such that $R_{mq} \sim U(0,.3)$, ensuring that overall allele frequencies in each population were maintained.
-We then used the methods described above to estimate $P({S_{jq}})$ and $E(N_{S})$ and their variances given $\gamma = 30$.
-For comparison, we also conducted 10,000 random draws of size $\gamma$ from each loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws.
-We likewise calculated the variance of $N_{S}$ directly across all sets of random draws.
+We then used the methods described above to estimate $P({S_{jq}})$, $E(N_{S})$, their variances, and 95% confidence and prediction intervals for $\gamma = $ 10, 20, ..., 100.
+For comparison, we also conducted 1000, 2000, ..., 10,000 random draws for each $\gamma$ from every loci in each population, then calculated $P({S_{jq}})$ and its variance for each locus empirically and $E(N_{S})$ by summing across all 100 loci for each set of draws.
 
-To compare our calculated estimates to the empirical simulations, we used the implementation of the Agresti-Coull [@agrestiApproximateBetterExact1998] method from the R package `binom` [@dorai-rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each empirical $P({S_{jq}})$.
-We also used normal approximation to calculate 95% confidence and prediction intervals for $E(N_{S})$ and $N_{S}$ using $\sigma_{E(n_{S})}^{2}$ and the variance directly observed from the simulations, respectively.
+To compare our calculated estimates to the empirical simulations, we used the implementation of the Agresti-Coull [@agrestiApproximateBetterExact1998] method from the R package `binom` [@dorai-rajBinomBinomialConfidence2022] to calculate 95% confidence intervals for each empirical $P({S_{jq}})$ derived from the simulations described above.
 
@@
 The methods described here for calculating $P({S_{j}})$ and $E(N_{S})$ performed well.
-Individual $P({S_{jq}})$ values for each locus were within the confidence intervals derived from the simulated values for 97 and 96% of loci from the $n = 100$ and $n = 1000$ populations, respectively (Figure 1).
-Note that variation along the generally correlated minor allele frequency/$P({S_{jq}})$ axes is due to variations in genotype frequencies in the simulated data for a given minor allele frequency.
-$P({S_{jq}})$ values estimated using equation 4 account for this adequately.
+Individual $P({S_{jq}})$ values for each locus and $\gamma$ from each population were within the 95% confidence intervals calculated from their respective simulations ~95% of the time, with no substantial bias across minor allele frequency or the number of simulations (Figures 1-2), although confidence intervals calculated from simulations with higher $\gamma$ values tended to contain $P({S_{j}})$ slightly more often (Figure S1).
+Note that variation along the generally correlated minor allele frequency/$P({S_{jq}})$ axes visible in Figure 1 due to variations in genotype frequencies in the simulated data for a given minor allele frequency.
+Given that the $P({S_{jq}})$ values estimated using equation 4 track the confidence intervals calculated from simulations, this is accounted for adequately.
 
-$E(N_{S})$ values estimated with equation 5 were similarly accurate and were within the 95% confidence intervals produced using the simulated $N_{S}$ for both population sizes (Figure 2).
-Likewise, the 95% prediction intervals calculated using $\sigma_{E(n_{S})}^{2}$ via equation 7 contained 96.1 and 95.3% of the simulated $N_{S}$ values for $n = 100$ and $n = 1000$, respectively (Figure 2).
+Calculations of the total number of segregating sites after rarefaction were likewise very similar to those observed via simulating random draws for any $\gamma$ or iteration counts.
+Specifically, $E(N_{S})$ values calculated via Equation 5 were very close to the mean number of segregating sites in each population after rarefaction (Figure 3, Figure S2).
+95% prediction intervals were generally very close to observed 95% quantiles from simulations across $\gamma$ and simulation counts but were consistatntly slightly overestimated on both ends.
+This is expected given that Equation 9 assumes normality; the actual distribution of $N_{S}$ values are slightly non-normal across samples.
+Specifically, the medians, but not the means, of the observed samples are therefore slightly above $E(N_{S})$ (Figure S2).
+Both point estimates of the number of segregating sites and 95% prediction intervals are similarly accurate across a wide range of both $\gamma$ values and simulation counts (Figure 4).
 
@@
 
-Figure 1: The expected probabilities $P({S_{j,a}})$ of observing a segregating site at a locus for loci with different minor allele frequencies.
-Probabilities are derived from Equation 4 for each locus *a* for population sizes of 100 (left) and 1000 (right), each rarefacted to $n = 30$.
-Each locus has a random, independent percentage of missing data which varies between 0 and 30.
-Points are colored depending on if they fell within a 95% confidence level based on 10,000 random re-samplings of each population to 30 individuals.
+Figure 1: The expected probabilities  of observing a segregating site at each locus ($P({S_{jq}})$) for loci with different minor allele frequencies.
+Probabilities are derived from Equation 4 for each locus for population sizes of 250 and 2500 rarefacted to either $N = 10$ or $N = 100$, corresponding to $\gamma = 10$ or $\gamma = 100$.
+Points vary from the trend lines due to the addition of missin data: each locus has a random, independent percentage of missing data which varies between 0 and 30.
+Points are colored depending on if they fell within a 95% confidence interval based on 10,000 simulated rarefaction trials of each population to each $N$, which is marked for each point with an error bar.
 
-Figure 2: The expected total number of segregating sites $E(N_{S})$ from populations of either 100 or 1,000 sampled to 30 individuals as calculated via Equation 5 vs that observed from 10,000 random re-samplings to 30 individuals.
-95% Confidence and prediction intervals for $E(N_{S})$ are shown, calculated according to Equations 8 and 9.
-Observed $N_{S}$ values from each re-sample are colored depending on if they fell within the estimated 95% prediction interval.
-The mean $N_{S}$ across all re-samples is marked with a triangle, colored depending on if it fell within the estimated 95% confidence interval.
+Figure 2: Trends in difference between the mathematically expected probability that a locus segregates after rarefaction $P({S_{jq}})$ and the observed probability of segregation following simulated rarefaction across different rarefaction sizes ($\gamma$), minor allele frequencies, and number of simulated rarefaction events.
+Loci are colored depending on if the expected probabilities fell within the observed 95% prediction intervals.
+
+Figure 3: The distribution of the total number of segregating sites ($N_{S}$) observed for 10,000 replicate simulated rarefaction trials to either $N = 10$ or $N = 100$ (corresponding to either $\gamma = 10$ or $\gamma = 100$, respectively) for starting population sizes of either $N = 250$ or $N = 2500$.
+The mathematically expected number of segregating sites ($E(N_{S})$) and 95% prediction intervals on that number are shown with solid yellow or dashed blue lines for each for each distribution.
+
+Figure 4: Trends in the distribution of the total number of segregating sites ($N_{S}$) observed following rarefaction trails across a range of rarefaction sizes ($\gamma$) and number of trials for starting population sizes of either $N = 250$ or $N = 2500$.
+Horizontal lines on each distribution note 95% quantile limits.
+The mathematically expected number of segregating sites ($E(N_{S})$) and 95% prediction intervals on that number are shown in yellow error bars to the right of each distribution.

### 2024-04-09 scripts/seg_sites_rarefaction.Rmd
changes prior to moving to doc for final

@@
 
+F1data$in_ci <- ifelse(F1data$in_ci, "Yes", "No")
+
 f1 <- ggplot(F1data[F1data$source == "Mathematical",],
              aes(x = af, y = p)) +
-  geom_point(aes(color = in_ci)) +
+  geom_point(aes(color = in_ci), size = 2.5) +
   geom_errorbar(aes(ymax = upper, ymin = lower)) +
@@
   scale_color_manual(values = colors) +
-  guides(color = guide_legend(title = "In CI")) +
-  ylab(bquote(italic(P*"("*S[jq]*")"))) +
+  guides(color = guide_legend(title = "In 95% CI")) +
+  ylab(bquote(italic(hat(P)*"("*S[jq]*")"))) +
   xlab("Minor Allele Frequency") +
@@
 
-# Figure 2: Distribution of Nseg vs Eseg for four runs
+# Figure 3: Distribution of Nseg vs Eseg for four runs
+Note that the label are flipped here: this produces figure 3.
 
@@
 
+K20_data <- which(ks == 20 & n_iters == 10000)
+F1data_K20 <- prep_f1(pseg1 = pseg1$K_20, 
+                       pseg2 = pseg2$K_20, 
+                       mpseg1 = res_files[[K20_data]]$mpseg1,
+                       mpseg2 = res_files[[K20_data]]$mpseg2,
+                       n1 = n1,
+                       n2 = n2)
+
 K10_data <- which(ks == 10 & n_iters == 10000)
@@
 
-K100_data <- which(ks == 100 & n_iters == 10000)
-F2data_K100 <- prep_f2(seg1 = res_files[[K100_data]]$seg1, 
-                       seg2 = res_files[[K100_data]]$seg2, 
-                       vseg1 = vseg1$K_100,
-                       vseg2 = vseg2$K_100,
-                       F1data = F1data_K100,
+K20_data <- which(ks == 20 & n_iters == 10000)
+F2data_K20 <- prep_f2(seg1 = res_files[[K20_data]]$seg1, 
+                       seg2 = res_files[[K20_data]]$seg2, 
+                       vseg1 = vseg1$K_20,
+                       vseg2 = vseg2$K_20,
+                       F1data = F1data_K20,
                        n1 = n1,
@@
 F2data_K10$sum_tab$K <- "gamma * ' = 10'"
-F2data_K100$sum_tab$K <- "gamma * ' = 100'"
+F2data_K20$sum_tab$K <- "gamma * ' = 20'"
 F2data_K10$all_mcmc_sums$K <- "gamma * ' = 10'"
-F2data_K100$all_mcmc_sums$K <- "gamma * ' = 100'"
+F2data_K20$all_mcmc_sums$K <- "gamma * ' = 20'"
 F2data_K10$pi_tab$K <- "gamma * ' = 10'"
-F2data_K100$pi_tab$K <- "gamma * ' = 100'"
+F2data_K20$pi_tab$K <- "gamma * ' = 20'"
 
-F2data_sums <- rbind(F2data_K10$sum_tab, F2data_K100$sum_tab)
+F2data_sums <- rbind(F2data_K10$sum_tab, F2data_K20$sum_tab)
 F2data_sums$pop <- gsub("n", "N", F2data_sums$pop)
@@
 
-F2data_mcmcs <- rbind(F2data_K10$all_mcmc_sums, F2data_K100$all_mcmc_sums)
+F2data_mcmcs <- rbind(F2data_K10$all_mcmc_sums, F2data_K20$all_mcmc_sums)
 F2data_mcmcs$pop <- gsub("n", "N", F2data_mcmcs$pop)
@@
 
-F2data_pi <- rbind(F2data_K10$pi_tab, F2data_K100$pi_tab)
+F2data_pi <- rbind(F2data_K10$pi_tab, F2data_K20$pi_tab)
 F2data_pi$pop <- gsub("n", "N", F2data_pi$pop)
@@
 
-f2 <- ggplot(F2data_mcmcs, aes(x = nseg)) +
-  geom_density() +
+F2data_mcmcs <- as.data.table(F2data_mcmcs)
+levs <- unique(F2data_mcmcs[,c("pop", "K")])
+F2data_densities <- vector("list", nrow(levs))
+
+for(i in 1:nrow(levs)){
+  d <- density(F2data_mcmcs[K == levs$K[i] & pop == levs$pop[i],]$nseg)
+  q <- quantile(F2data_mcmcs[K == levs$K[i] & pop == levs$pop[i],]$nseg, c(0.025,0.975))
+  F2data_densities[[i]] <- data.table(x = d$x, y = d$y, K = levs$K[i], pop = levs$pop[i])
+  F2data_densities[[i]]$quant <- factor(findInterval(F2data_densities[[i]]$x, q))
+}
+
+F2data_densities <- rbindlist(F2data_densities)
+
+f2 <- ggplot(F2data_densities, aes(x = x, y = y)) +
+  geom_line() +
+  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), show.legend = FALSE) +
   geom_vline(data = F2data_sums, aes(xintercept = nseg), color = colors[1]) +
@@
   scale_y_continuous(expand = c(0,0)) +
+  scale_fill_manual(values = c("#004488", "#ffffff", "#004488")) +
   theme(axis.text = element_text(size = t.sizes[2]),
@@
 
-ggsave("Figure2.pdf", f2, height = 10, width = 15)
+ggsave("Figure3.pdf", f2, height = 10, width = 15)
 
@@
 
-# Figure 3: var vs difference between estimate and mean of simulations for mcmc and k
+# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k
+Note that the labels are flipped here, this produces figure 2
 
@@
   # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
+  theme(axis.text = element_text(size = t.sizes[2] - 2),
+        axis.title = element_text(size = t.sizes[1] - 2),
+        strip.text = element_text(size = t.sizes[1] - 2),
         strip.background = element_blank(),
-        legend.title = element_text(size = t.sizes[1]),
-        legend.text = element_text(size = t.sizes[2])) +
+        legend.title = element_text(size = t.sizes[1] - 2),
+        legend.text = element_text(size = t.sizes[2] - 2)) +
   xlab(parse(text = "gamma")) +
@@
 
-# Figure S2: maf vs error rate
-ps_all_aggregate_for_af <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(af, iters, pop)]
-FS2 <- ggplot(ps_all_aggregate_for_af, aes(x = af, y = in_ci.mathematical*100)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(iters~pop) +
-  theme_bw() +
-  # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
-  # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
-        strip.background = element_blank(),
-        legend.title = element_text(size = t.sizes[1]),
-        legend.text = element_text(size = t.sizes[2])) +
-  xlab("Minor Allele Frequency") +
-  ylab("% estimates in CI")
+save_plot("FigureS1.pdf", FS1, base_height = 10, base_width = 15)
 
-f3a <- ggplot(ps_all_merge, aes(x = K, y = diff, color = in_ci.mathematical)) +
-  geom_point() +
+# # Figure S2: maf vs error rate
+# ps_all_aggregate_for_af <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(af, iters, pop)]
+# FS2 <- ggplot(ps_all_aggregate_for_af, aes(x = af, y = in_ci.mathematical*100)) +
+#   geom_point() +
+#   geom_smooth(method = "lm", se = FALSE) +
+#   facet_grid(iters~pop) +
+#   theme_bw() +
+#   # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
+#   # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
+#   theme(axis.text = element_text(size = t.sizes[2]),
+#         axis.title = element_text(size = t.sizes[1]),
+#         strip.text = element_text(size = t.sizes[1]),
+#         strip.background = element_blank(),
+#         legend.title = element_text(size = t.sizes[1]),
+#         legend.text = element_text(size = t.sizes[2])) +
+#   xlab("Minor Allele Frequency") +
+#   ylab("% estimates in CI")
+
+ps_all_merge[,in_ci.mathematical := ifelse(in_ci.mathematical, "Yes", "No")]
+ps_all_merge[,in_ci.mathematical := factor(in_ci.mathematical, c("Yes", "No"))]
+f3a <- ggplot(ps_all_merge, aes(x = K, y = diff, color = in_ci.mathematical, group = in_ci.mathematical)) +
+  geom_point(position = position_dodge(4)) +
   theme_bw() +
@@
   xlab(parse(text = "gamma")) +
-  ylab(bquote(italic(P*"("*S[jq]*")"[E]) - italic(P*"("*S[jq]*")"[O]))) +
+  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")"))) +
   guides(color = guide_legend(title = "In 95% CI"))
 
-f3b <- ggplot(ps_all_merge, aes(x = af, y = diff, color = in_ci.mathematical)) +
-  geom_point() +
+f3b <- ggplot(mapping = aes(x = af, y = diff, color = in_ci.mathematical)) +
+  geom_point(data = ps_all_merge[in_ci.mathematical == "Yes",]) +
+  geom_point(data = ps_all_merge[in_ci.mathematical == "No",]) +
   theme_bw() +
@@
         axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
+        strip.text = element_blank(),
         axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
@@
   xlab("Minor Allele Frequency") +
-  ylab(bquote(italic(P*"("*S[jq]*")"[E]) - italic(P*"("*S[jq]*")"[O])))
+  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")")))
 
-f3c <- ggplot(ps_all_merge, aes(x = iters, y = diff, color = in_ci.mathematical)) +
-  geom_point() +
+f3c <- ggplot(ps_all_merge, aes(x = iters, y = diff, color = in_ci.mathematical, group = in_ci.mathematical)) +
+  geom_point(position = position_dodge(400)) +
   theme_bw() +
@@
         axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
+        strip.text = element_blank(),
         axis.text.x = element_text(size = t.sizes[2], 
@@
   xlab("Number of Simulated Draws") +
-  ylab(bquote(italic(P*"("*S[jq]*")"[E]) - italic(P*"("*S[jq]*")"[O])))
+  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")")))
 
@@
 
-save_plot("Figure3.pdf", f3, base_height = 11, base_width = 15)
+save_plot("Figure2.pdf", f3, base_height = 11, base_width = 15)
 ```
@@
   geom_errorbar(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper")]),
-            aes(x = K, ymin = lower, ymax = upper), color = "#DDAA33", size = 1, 
-            position = position_nudge(x = 3.5),
+            aes(x = K + 3.7, ymin = lower, ymax = upper), color = "#DDAA33", size = 1,
             inherit.aes = FALSE, width = 2) +
-  stat_ydensity(aes(group = K), draw_quantiles = c(.025, .975), scale = "width", bw = .4,
+  geom_point(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper", "ENs")]),
+             aes(x = K + 3.7, y = ENs), color = "#DDAA33", size = 7, shape = 45,
+            inherit.aes = FALSE) +
+  stat_ydensity(aes(group = K), draw_quantiles = c(.025, .5, .975), scale = "width", bw = .4,
                 fill = NA, width = 6) +
   facet_wrap(~pop) +
+  scale_y_continuous(breaks = seq(40, 100, by = 10), limits = c(40, 100)) +
   theme_bw() + 
@@
         legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(E*"("*N[S]*")") - italic(N[S,observed]))) +
+  ylab(bquote(italic(N[S]))) +
   xlab(bquote(gamma))
@@
   geom_errorbar(data = unique(ns_all[K == 50,c("pop", "K", "iters", "i", "lower", "upper")]),
-            aes(x = iters, ymin = lower, ymax = upper), color = "#DDAA33", size = 1, 
-            position = position_nudge(x = 350),
+            aes(x = iters + 370, ymin = lower, ymax = upper), color = "#DDAA33", size = 1,
             inherit.aes = FALSE, width = 200) +
-  stat_ydensity(aes(group = iters), draw_quantiles = c(.025, .975), scale = "width", bw = .4,
+  geom_point(data = unique(ns_all[K == 50, c("pop", "K", "iters", "i", "lower", "upper", "ENs")]),
+             aes(x = iters + 370, y = ENs), color = "#DDAA33", size = 7, shape = 45,
+            inherit.aes = FALSE) +
+  stat_ydensity(aes(group = iters), draw_quantiles = c(.025, .5, .975), scale = "width", bw = .4,
                 fill = NA, width = 600) +
@@
   theme_bw() + 
+  scale_y_continuous(breaks = seq(40, 100, by = 5), limits = c(80, 100)) +
   scale_x_continuous(breaks = mcmcs) +
@@
         legend.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
+        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
+        strip.text = element_blank(),
         strip.background = element_blank(),
         legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(E*"("*N[S]*")") - italic(N[S,observed]))) +
+  ylab(bquote(italic(N[S]))) +
   xlab("Number of Simulated Draws")
 
-```
+f4 <- plot_grid(f4a, f4b, ncol = 1, axis = "l", align = "l")
+save_plot("Figure4.pdf", f4, base_height = 11, base_width = 15)
 
-```{r}
+# S2
+ns_all_agg <- ns_all[,lapply(.SD,mean), .SDcols = "nseg", by = c("ENs", "pop", "K", "iters", "i")]
+levs <- paste0("Sims = ", sort(unique(ns_all_agg$iters)))
+ns_all_agg$iters <- factor(paste0("Sims = ", ns_all_agg$iters), levels = levs)
+ns_all_agg$pop <- gsub("n", "N", ns_all_agg$pop)
+ns_all_agg$stat <- "Mean"
 
-#===============plot all of the observed values vs the estimated value for the sums=============
-# plot the estimated prediction interval vs the observations and the confint from the simultations
+ns_all_agg_med <- ns_all[,lapply(.SD,median), .SDcols = "nseg", by = c("ENs", "pop", "K", "iters", "i")]
+levs <- paste0("Sims = ", sort(unique(ns_all_agg_med$iters)))
+ns_all_agg_med$iters <- factor(paste0("Sims = ", ns_all_agg_med$iters), levels = levs)
+ns_all_agg_med$pop <- gsub("n", "N", ns_all_agg_med$pop)
+ns_all_agg_med$stat <- "Median"
 
-#==========others==================
+ns_all_agg <- rbind(ns_all_agg, ns_all_agg_med)
+FS2 <- ggplot(ns_all_agg, aes(x = K, y = ENs - nseg, color = pop)) +
+  geom_point() +
+  facet_grid(stat~iters) +
+  khroma::scale_color_highcontrast() +
+  geom_hline(yintercept = 0) +
+  theme_bw() +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        legend.title = element_text(size = t.sizes[1]),
+        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
+        strip.background = element_blank(),
+        strip.text = element_text(size = t.sizes[1]),
+        legend.text = element_text(size = t.sizes[2])) +
+  ylab(bquote(italic(E(N[S])) * " - " * italic(N[S,observed]))) +
+  xlab(bquote(gamma)) +
+  guides(color = guide_legend(title = "Population"))
 
-# variances, per locus
-vars <- rbind(cbind.data.frame(estimated = vseg1, 
-                               observed = matrixStats::rowVars(seg1),
-                               pop = paste0("n = ", n1/2),
-                               af = allele_frequencies),
-              cbind.data.frame(estimated = vseg2, 
-                               observed = matrixStats::rowVars(seg2),
-                               pop = paste0("n = ", n2/2),
-                               af = allele_frequencies))
-ggplot(vars, aes(x = observed, y = estimated, color = af)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  facet_wrap(~pop) +
-  theme_bw() +
-  khroma::scale_color_batlow()
+save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 15)
 
-# variances, overall
-overall_vars <- rbind(cbind.data.frame(estimated = sum(vseg1),
-                                       observed = var(msum1),
-                                       pop = paste0("n = ", n1/2)),
-                      cbind.data.frame(estimated = sum(vseg2),
-                                       observed = var(msum2),
-                                       pop = paste0("n = ", n2/2)))
 ```

### 2024-04-09 scripts/seg_sites_rarefaction.Rmd
supplementary data and snpR per-locus for SI

@@
 # Rarefact via simulations
-Note: this chunck should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunck.
+Note: this chunk should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunk.
 `q` should be passed as an argument for the iteration, as should `outfile`.
@@
 ```{r cluster_MCMCs, eval=FALSE}
-set.seed(seed + q) # reset seed for cluster
+# set.seed(seed + q) # reset seed for cluster
 
@@
 
-# draw for each k/mcmc count combination
-combs <- expand.grid(k, mcmcs)
-colnames(combs) <- c("k", "iters")
-
-tmcmcs <- combs$iters[q]
-tk <- combs$k[q]
-
-seg1 <- seg2 <- matrix(0, nal, tmcmcs)
-for(i in 1:tmcmcs){
-  for(j in 1:nal){
-    seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
-    seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
-  }
-}
-
-mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
-                               method = "agresti-coull")[,c("mean", "lower", "upper")]
-mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
-                               method = "agresti-coull")[,c("mean", "lower", "upper")]
-
-colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
-
-res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
-
-saveRDS(res, paste0(outfile, "_", q, ".RDS"))
+# # draw for each k/mcmc count combination
+# combs <- expand.grid(k, mcmcs)
+# colnames(combs) <- c("k", "iters")
+# 
+# tmcmcs <- combs$iters[q]
+# tk <- combs$k[q]
+# 
+# seg1 <- seg2 <- matrix(0, nal, tmcmcs)
+# for(i in 1:tmcmcs){
+#   for(j in 1:nal){
+#     seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
+#     seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
+#   }
+# }
+# 
+# mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
+#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
+# mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
+#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
+# 
+# colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
+# 
+# res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
+# 
+# saveRDS(res, paste0(outfile, "_", q, ".RDS"))
 
@@
 ```
+
+#========real data example, mathematically============
+
+```{r}
+library(snpR)
+set.seed(12321)
+d <- readRDS("../data/monarchs_non_strict.RDS")
+d <- d[sample(nrow(d), 1000),] # subsample 1k loci
+
+# remove smaller than 10, although it works fine including stuff like SAI with only 4
+sample.meta(d)$pop[sample.meta(d)$pop %in% c("ENA", "WNA")] <- "NAM"
+ps <- summarize_facets(d, "pop")$pop
+d <- d[pop = -c(names(ps)[which(ps < 10)])]
+
+d <- calc_seg_sites(d, "pop", g = 0) # set gamma to the smallest N per loci
+seg_res_mon <- get.snpR.stats(d, "pop", "seg_sites") # follows biogeographic history
+```
+
+#=========real data example, via rarefaction==========
+```{r}
+# setup
+sn <- format_snps(d, "sn", interpolate = FALSE)
+sn <- sn[,-c(1:7)]
+tmcmcs <- 100
+
+pops <- substr(colnames(sn), 1, 3)
+upops <- unique(pops)
+mon_seg <- array(0, c(nrow(sn), tmcmcs, length(unique(pops))))
+
+# do the rarefaction
+for(i in 1:tmcmcs){
+  cat("\nIter:", i)
+  for(j in 1:nrow(sn)){
+    # cat("\n\t", j)
+    tg <- min(tapply(unlist(sn[j,]), pops, function(x) sum(!is.na(x)))) # set gamma for this locus
+    for(k in 1:length(upops)){
+      mon_seg[j,i,k] <- rarefact_mcmc(sn[j,which(pops == upops[k])], k = tg) # rarefact this locus in this pop
+    }
+  }
+}
+
+# summarize results
+mon_rare_res <- matrix(0, tmcmcs, length(upops))
+for(k in 1:length(upops)){
+  mon_rare_res[,k] <- colSums(mon_seg[,,k])
+}
+colnames(mon_rare_res) <- upops
+mon_rare_res <- as.data.table(mon_rare_res)
+mon_rare_res$i <- 1:tmcmcs
+mon_rare_res <- melt(mon_rare_res, id.vars = "i", variable.name = "pop", value = "Ns")
+
+ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
+  geom_boxplot() +
+  geom_point(alpha = .2, position = position_jitter(height = 0, width = .2)) +
+  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red") +
+  theme_bw() +
+  ylab(bquote(italic(N[s]))) +
+  xlab("Population")
+
+```

### 2024-04-22 manuscript/SI_temp.Rmd
files pre-publication for movement to christeLab repo

---
title: "Untitled"
output:
  word_document: default
  html_document: default
date: "2024-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

Install:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("hemstrow/snpR")
```

load
```{r}
library(snpR)
```

download

```{r}
# download the metadata
meta <- read.table(url("https://raw.githubusercontent.com/hemstrow/seg_sites_rarefaction/main/data/example_metadata.txt"), 
                   header = TRUE)

# download the vcf
download.file("https://raw.githubusercontent.com/hemstrow/seg_sites_rarefaction/main/data/example_vcf.vcf", 
              destfile = "example_vcf.vcf")
```

read in
```{r}
monarchs <- read_vcf("example_vcf.vcf", sample.meta = meta)
```

types
```{r}
summarize_facets(monarchs, "pop")
monarchs <- calc_seg_sites(monarchs, "pop", g = 0)
get.snpR.stats(monarchs, "pop", "seg_sites")$weighted.means

monarchs <- calc_seg_sites(monarchs, "pop", g = -1)
get.snpR.stats(monarchs, "pop", "seg_sites")$weighted.means

monarchs <- calc_seg_sites(monarchs, "pop", g = 10)
get.snpR.stats(monarchs, "pop", "seg_sites")$weighted.means

monarchs <- calc_seg_sites(monarchs, "pop", g = 0)
head(get.snpR.stats(monarchs, "pop", "seg_sites")$single, 10)
```

### 2024-04-22 manuscript/ms.tex
files pre-publication for movement to christeLab repo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setsansfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage[left]{lineno}
\linenumbers
\usepackage{setspace}\doublespacing
\usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
\usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
\usepackage[round]{natbib}
\usepackage{amsmath}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={manuscript},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{manuscript}
\author{}
\date{\vspace{-2.5em}31 October 2023}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{probability-of-observing-segregating-loci-via-rarefaction}{%
\subsection{Probability of observing segregating loci via
rarefaction}\label{probability-of-observing-segregating-loci-via-rarefaction}}

Rarefaction can be used to estimate the probability of observing a
segregating site at a specific locus using much the same framework used
to calculate allelic richness. In brief, allelic richness (or the
expected number of distinct alleles expected at a given locus under a
common sample size \(g\) across populations) can be estimated for a
given population \(j\) by summing the probability of observing each
\(i\) of \(m\) unique alleles using the counts of those alleles in the
population \(N_{ij}\) and the total sample size in that population
\(N_{j}\). This is done by comparing the number of possible ways to draw
\(g\) gene copies without sampling allele \(i\)
(\(\binom{N_{j} - N_{ij}}{g}\)) to the total number of possible
combinations of gene copies that can be drawn (\(\binom{N_{j}}{g}\));
the inverse of this
(\(1 - \frac{\binom{N_{j} - N_{ij}}{g}}{\binom{N_{j}}{g}}\)) is
therefore the probability of observing allele \(i\) in population
\(N_{j}\), and the sum of this value across all \(m\) alleles gives the
expected number of alleles observed at a locus in population \(j\),
\(\alpha_{g}^{j}\) (\textbf{hurlbertNonconceptSpeciesDiversity1971?};
\textbf{kalinowskiCountingAllelesRarefaction2004?}):

\begin{equation}
\alpha_{g}^{j} = \sum_{i = 1}^{m} 1 - \frac{\binom{N_{j} - N_{ij}}{g}}{\binom{N_{j}}{g}}
\end{equation}

The expected number of segregating loci in a population for a draw of
\(g\) gene copies can be derived similarly. For a locus \(i\) to be
segregating in population \(j\), all alleles drawn across all gene
copies must be identical. If alleles are independent at each locus (the
locus is at Hardy-Weinburg Equlibrium, HWE) and \(N\) is infinite, the
probability (\(P({S_{j}})\)) of observing a segregating site at a locus
is the inverse of the probability of drawing only one allele in \(g\)
draws with replacement: \[P({S_{j}}) = 1 - \sum_{i = 1}^{m} f_{ij}^{g}\]

where \(f_{ij}\) is the allele frequency of allele \(i\) in population
\(j\). However, in finite samples draws are conducted with replacement,
and so binomial coefficients must instead be used to determine the
probability of drawing only a specific allele:
\[P({S_{j}}) = 1 - \sum_{i = 1}^{m} \frac{\binom{N_{ij}}{g}}{\binom{N_{j}}{g}}\]

However, HWE is often not a desirable assumption to make. Even if
filtering is employed to remove loci which do not conform to HWE, the
degree of conformity, and thus the degree of statistical bias in
estimating \(P({S_{j}})\), typically varies somewhat between
populations. For example, in a sample of 100 genotypes with a minor
allele frequency of 0.05, only five minor alleles are expected and two
out of three possible combinations of minor homozygotes and
heterozygotes that produce that frequency will not deviate from HWE at
\(\alpha = 0.05\) according to an exact test (Wigginton et al., 2005).
However, re-sampling these to, say, ten genoytpes should will produce
quite different \(P(S_{j})\) (roughly 0.7, 0.6, and 0.5, for purely
heterozygotes, one homozygote and three heterozygotes, and two
homozygotes and one heterozygote, respectively) as we will see below.

To remedy these problems, I propose the following estimator of
\(P({S_{j}})\):
\[P({S_{j}}) = 1 - \sum_{k = 1}^{h}\frac{\binom{n_{kj}}{\gamma}}{\binom{n_{j}}{\gamma}}\]
where the \(P({S_{j}})\) is given by the probability of exclusively
drawing any \(k\) of \(h\) possible homozygote genotypes in population
\(j\) given \(\gamma\) independent sampled \emph{genotypes} (not
\emph{gene copies}) from the pool of observed genotypes. Here,
\(n_{kj}\) is the number of observed homozygote genotypes of type \(k\)
in population \(j\) and \(n_{j}\) is the total number of observed
genotypes of all types. Note that \(n_{j}\) and \(\gamma\) will be half
the value of their equivalents \(N_{j}\) and \(g\) for diploid species,
one third for triploids, and so on.

Interestingly, this method, like the richness method and related private
allele rarefaction approaches can smoothly account for varying amounts
of missing data at specific loci in different populations by varying
\(g\) or \(\gamma\) across loci. Both can be set to one less than the
smallest observed \(N_{j}\) or \(n_{j}\), the highest values at which
rarefaction can be applied within a population, across all populations
after accounting for missing data, and can thus vary across loci without
bias. Setting either value to \(N_{j}\) or \(n_{j}\) will instead return
the observed allele diversity or segregating site status, respectively.

This is particularly useful given that \(E(N_{S})\) or the expected
total number of segregating sites, is often of specific interest as a
measure of genetic diversity when comparing populations. Given that the
expected number of segregating sites at a specific locus \(q\) in
population \(j\), \(E(N_{S_{jq}})\), is equal to \(P({S_{jq}})\),
\(E(N_{S})\) can be calculated by summing \(P({S_{jq}})\) across all
\(Q\) loci: \[E(N_{S}) = \sum_{q = 1}^{Q} P({S_{jq}})\] with \(\gamma\)
set accordingly for each locus. In this case,
\(0 \le E(N_{S_{jq}}) \le 1\) for all loci (and thus
\(0 \le E(N_{S}) \le Q\)).

Usefully, under this framework each locus represents a single Bernoulli
trial in which it can be observed to be segregating or not with
probability \(P({S_{jq}})\). As such, the variance of \(P({S_{jq}})\)
for each locus is given by
\[\sigma_{P(S_{jq})}^{2} = P(S_{jq}) \times (1 - P(S_{jq}))\] and, if
each locus is independent, the variance of \(E(N_{S})\) is equal to the
sum of \(\sigma_{P(S_{jq})}^{2}\) across all loci:
\[\sigma_{E(n_{S})}^{2} = \sum_{q = 1}^{Q}\sigma_{P(S_{jq})}^{2}\]

Confidence and prediction intervals can then be derived using standard
approaches for the sum of random, independent Bernoulli trials. When
\(Q\) is large, for example, the distribution of \(E(N_{S})\) should be
approach normal and confidence and prediction intervals can be derived
using standard normal approximation using the equations
\[CI_{N_{S}} = E(N_{S}) \pm Z\sqrt{\frac{\sigma_{E(n_{S})}^{2}}{Q}}\]
and \[PI_{N_{S}} = E(N_{S}) \pm Z\sqrt{\sigma_{E(n_{S})}^{2}(1+(1/Q))}\]
where \(Z\) is given by the normal quantile function
\(Z = Q_{X}(1-\alpha)\) with \(\mu = 0\) and \(\sigma = 1\) for a
desired confidence level \(\alpha\).

\hypertarget{emperical-validation}{%
\subsection{Emperical Validation}\label{emperical-validation}}

To validate equations 4 and 5, I simulated genotypic data for two
populations with sizes 100 and 1000, each with 100 bi-allelic loci with
minor allele frequencies spaced equally between 0.01 and 0.1. I added
missing data to each population assigning each locus a missing data rate
\(R_{mq}\) from a uniform distribution such that
\(R_{mq} \sim U(0,.3)\), ensuring that overall allele frequencies in
each population were maintained. I then used the methods described above
to estimate \(P({S_{jq}})\) and \(E(N_{S})\) and their variances given
\(\gamma = 30\). For comparison, I also conducted 10,000 random draws of
size \(\gamma\) from each loci in each population, then calculated
\(P({S_{jq}})\) and its variance for each locus empirically and
\(E(N_{S})\) by summing across all 100 loci for each set of draws. I
likewise calculated the variance of \(N_{S}\) directly across all sets
of random draws.

To compare my calculated estimates to the empirical simulations, I used
the ``exact'' method from the R package \texttt{binom}
(\textbf{rajBinomBinomialConfidence2022?}) to calculate 95\% confidence
intervals for each empirical \(P({S_{jq}})\). I also used normal
approximation to calculate 95\% confidence and prediction intervals for
\(E(N_{S})\) and \(N_{S}\) using \(\sigma_{E(n_{S})}^{2}\) and the
variance directly observed from the simulations, respectively.

\hypertarget{results}{%
\section{Results}\label{results}}

The methods described here for calculating \(P({S_{j}})\) and
\(E(N_{S})\) performed well. Individual \(P({S_{jq}})\) values for each
locus were within the confidence intervals derived from the simulated
values for 97 and 96\% of loci from the \(n = 100\) and \(n = 1000\)
populations, respectively (Figure 1). Note that variation along the
generally correlated minor allele frequency/\(P({S_{jq}})\) axes is due
to variations in genotype frequencies in the simulated data for a given
minor allele frequency. \(P({S_{jq}})\) values estimated using equation
X account for this adequately.

\(E(N_{S})\) values estimated with equation X were similarly accurate
and were within the 95\% confidence intervals produced using the
simulated \(N_{S}\) for both population sizes. Likewise, the 95\%
prediction intervals calculated using \(\sigma_{E(n_{S})}^{2}\) via
equation X contained 96.1 and 95.3\% of the simulated \(N_{S}\) values
for \(n = 100\) and \(n = 1000\), respectively.

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Wigginton2005}{}}%
Wigginton, J. E., Cutler, D. J., \& Abecasis, G. R. (2005). A {Note} on
{Exact} {Tests} of {Hardy}-{Weinberg} {Equilibrium}. \emph{The American
Journal of Human Genetics}, \emph{76}(5), 887--893.
https://doi.org/\url{https://doi.org/10.1086/429864}

\end{CSLReferences}

\end{document}

### 2024-04-22 scripts/seg_sites_rarefaction.Rmd
files pre-publication for movement to christeLab repo

@@
 
-```{r cluster_MCMCs, eval=FALSE}
+```{r cluster_MCMCs}
 # set.seed(seed + q) # reset seed for cluster
@@
 ```{r}
-res_files <- list.files("../data/", "\\.RDS", full.names = TRUE)
+res_files <- list.files("../data/", "mcmc.+\\.RDS", full.names = TRUE)
 res_files <- lapply(res_files, readRDS)
@@
 prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
-  #========plot the emperical CIs around every loci vs the predicted value==============
-  # plotting the emperical CIs vs the estiamted mean
+  #========plot the empirical CIs around every loci vs the predicted value==============
+  # plotting the empirical CIs vs the estimated mean
   pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
@@
 
-# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k
+# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k; Figure S1
 Note that the labels are flipped here, this produces figure 2
@@
 
+  # per snp
   ps_prep$K <- tk
@@
 
+  ## counts seg per snp
+  cn1 <- res_files[[i]]$seg1
+  cn1 <- rowSums(cn1)
+  cn2 <- res_files[[i]]$seg2
+  cn2 <- rowSums(cn2)
+
+  cn <- rbind(data.frame(pop = paste0("N = ", n1/2), af = allele_frequencies, n_obs_seg = cn1),
+              data.frame(pop = paste0("N = ", n2/2), af = allele_frequencies, n_obs_seg = cn2))
+  ps_prep <- merge(ps_prep, cn, by = c("pop", "af"), sort = FALSE)
+
+  # total
   ns_prep$all_mcmc_sums$K <- tk
@@
 
-ps_all_m <- ps_all[source == "Mathematical",]
+ps_all_m <- ps_all[source == "Mathematical",] # note: all of the CIs in this refer to: is the mathematical expectation in the observed CI.
 ps_all_s <- ps_all[source == "Simulation",]
@@
 ps_all_merge[,diff := p.mathematical - p.simulated]
+
+# note: this code does the reverse: is the observed number of seg sites for a locus a set of sims within the PI expected based on the expected prob
+ps_all_merge[,pi_lower := qbinom(.05, iters, p.mathematical)]
+ps_all_merge[,pi_upper := qbinom(.95, iters, p.mathematical)]
+ps_all_merge[,in_pi := n_obs_seg.simulated >= pi_lower & n_obs_seg.simulated <= pi_upper]
 
@@
 
-save_plot("FigureS1.pdf", FS1, base_height = 10, base_width = 15)
+save_plot("FigureS1.pdf", FS1, base_height = 15, base_width = 15)
+save_plot("FigureS1.jpg", FS1, base_height = 15, base_width = 15)
 
@@
 ```
-# Figure 4 Nseg vs parameters
+# Figure 4 Nseg vs parameters; Figure S2
 
@@
 
-save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 15)
+save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 21)
+save_plot("FigureS2.jpg", FS2, base_height = 10, base_width = 21)
 
@@
 seg_res_mon <- get.snpR.stats(d, "pop", "seg_sites") # follows biogeographic history
+
+# save this as a VCF for the supplementary material example
+format_snps(d, "vcf", outfile = "../data/example_vcf.vcf", chr = "CHROM")
+fwrite(sample.meta(d), "../data/example_metadata.txt", col.names = TRUE, row.names = FALSE, sep = "\t")
 ```
@@
 
-ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
+FS3 <- ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
   geom_boxplot() +
   geom_point(alpha = .2, position = position_jitter(height = 0, width = .2)) +
-  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red") +
+  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red", size = 4) +
   theme_bw() +
   ylab(bquote(italic(N[s]))) +
-  xlab("Population")
+  xlab("Population") +
+  theme(axis.text = element_text(size = t.sizes[2]),
+        axis.title = element_text(size = t.sizes[1]),
+        legend.title = element_text(size = t.sizes[1]))
 
+save_plot("FigureS3.pdf", FS3, base_width = 15, base_height = 11)
+save_plot("FigureS3.jpg", FS3, base_width = 15, base_height = 11)
 ```

### 2024-09-30 scripts/Figure5.Rmd
pre-resubmission commit

---
title: "Figure5"
output: html_document
date: "2024-08-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE);library(snpR); library(ggplot2)
```

# Read in, phase, simulate data

First, read and phase:

```{r}
d <- readRDS("../data/monarch_nomaf.RDS")

chrs <- table(snp.meta(d)$group)
bad_chrs <- names(chrs)[which(chrs == 1)]
bad_snps <- which(snp.meta(d)$group %in% bad_chrs)
d <- import.snpR.data(genotypes(d)[-bad_snps,], snp.meta(d)[-bad_snps,], sample.meta(d))

format_snps(d, "vcf", chr = "group", outfile = "../data/monarchs.vcf")
cmd <- "java -Xmx4g -jar ~/bin/beagle.22Jul22.46e.jar gt=../data/monarchs.vcf out=../data/monarchs_phased"
system(cmd)
system("gunzip ../data/monarchs_phased.vcf.gz")
```

Next, read in the phased data and simulate randomly mating and outbred pops from it.

```{r}
dv <- data.table::fread("../data/monarchs_phased.vcf")
pops <- colnames(dv)[-c(1:9)]
pops <- substr(pops, 1, 3)

meta <- dv[,1:2]
dv <- dv[,-c(1:9), with = FALSE]

genotypes <- list(nam = dv[,which(pops == "NAM"), with = FALSE],
                  qld = dv[,which(pops == "QLD"), with = FALSE])
genotypes <- lapply(genotypes, function(x){
  x1 <- matrix(as.numeric(substr(as.matrix(x), 1, 1)), nrow = nrow(x))
  x2 <- matrix(as.numeric(substr(as.matrix(x), 3, 3)), nrow = nrow(x))
  return(cbind(x1, x2)[,order(c(1:ncol(x1), 1:ncol(x2)))])
})

effects <- rep(0, nrow(dv))
effects <- cbind(effects, effects)

target_size <- 50
chr_lengths <- tapply(meta$POS, meta$`#CHROM`, max)
colnames(meta) <- c("chr", "position")
set.seed(123343)

tno <- simulate_populations(genotypes, meta = meta, effects = effects, h = .5, gens = 2, chr_length = chr_lengths,
                            growth_function = function(n) BL_growth(n, 2),
                            var_theta = 0,
                            K_thin_post_surv = target_size,
                            thin = FALSE,
                            thin_fixed = FALSE, 
                            sampling_point = "offspring",
                            starting_surv_opt = c(0, 0),
                            migration = migration,
                            survival_function = function(phenotypes, ...) rep(1, length(phenotypes)),
                            selection_shift_function = function(opt, ...) opt,
                            verbose = TRUE)

tno2 <- simulate_populations(tno$genotypes[1], meta = tno$meta, effects = tno$effects[,1, drop = FALSE],
                             h = .5, gens = 5, chr_length = chr_lengths,
                             growth_function = function(n) BL_growth(n, 2),
                             var_theta = 0,
                             K_thin_post_surv = target_size,
                             thin = FALSE,
                             thin_fixed = FALSE, 
                             sampling_point = "offspring",
                             starting_surv_opt = c(0),
                             survival_function = function(phenotypes, ...) rep(1, length(phenotypes)),
                             selection_shift_function = function(opt, ...) opt,
                             verbose = TRUE)

genos <- cbind(genotypes[[1]], genotypes[[2]], tno$genotypes[[1]], tno$genotypes[[2]], tno2$genotypes[[1]])
genos <- GeneArchEst::convert_2_to_1_column(genos)
genos <- t(genos)
d <- import.snpR.data(as.data.frame(genos), snp.meta = tno$meta, 
                      sample.meta = data.frame(pop = c(rep("NAM", ncol(genotypes[[1]])/2),
                                                       rep("QLD", ncol(genotypes[[2]])/2), 
                                                       rep("OutbredA", ncol(tno$genotypes[[1]])/2),
                                                       rep("OutbredB", ncol(tno$genotypes[[2]])/2),
                                                       rep("Random", ncol(tno2$genotypes[[1]])/2))))
```

# Calculate summary stats

```{r}
d <- d[pop = c("NAM", "OutbredA", "Random")] # keeping these
d <- filter_snps(d, non_poly = TRUE)
set.seed (3218434)
d <- d[sample(nrow(d), 10000, FALSE),]
gc();

d <- calc_fis(d, "pop")
get.snpR.stats(d, "pop", "fis")$weighted.means

n <- 25 # sample size for rarefaction

# calc seg sites, watterson's theta, fis, projection S
d <- calc_seg_sites(d, "pop", g = n)
d <- calc_tajimas_d(d, "pop", global = TRUE)
d <- calc_fis(d, "pop")
res <- get.snpR.stats(d, "pop", c("tsd", "seg_sites", "fis"))$weighted.means
sfs1 <- calc_sfs(d, "pop", "NAM", projection = n*2)
sfs2 <- calc_sfs(d, "pop", "OutbredA", projection = n*2)
sfs3 <- calc_sfs(d, "pop", "Random", projection = n*2)
```

# Bootstrap rarefact

```{r}
one_sub <- function(d, n){
  d_sub <- d[,c(sample(which(sample.meta(d)$pop == "NAM"), n), 
                sample(which(sample.meta(d)$pop == "OutbredA"), n), 
                sample(which(sample.meta(d)$pop == "Random"), n))]
  d_sub <- filter_snps(d_sub, non_poly = TRUE, verbose = FALSE)
  d_sub <- calc_seg_sites(d_sub, "pop")
  res_sub <- get.snpR.stats(d_sub, "pop", c("tsd", "seg_sites"))$weighted.means

  return(res_sub)
}

boots <- 1000
sub_res <- vector("list", boots)

cl <- parallel::makePSOCKcluster(5)
doParallel::registerDoParallel(cl)

pboot <- split(1:boots, rep(1:5, length.out = length(1:boots), each = ceiling(length(1:boots)/5)))

sub_res <- foreach(q = 1:5, .packages = "snpR") %dopar% {
  set.seed(3218214 + q)

  ti <- pboot[[q]]
  tres <- vector("list", length(ti))
  for(i in 1:length(ti)){
    tres[[i]] <- one_sub(d, n)
    tres[[i]]$iter <- i
  }
  tres <- rbindlist(tres)
  tres
}
parallel::stopCluster(cl)

sub_res <- data.table::rbindlist(sub_res)
sub_res[,fis := res$weighted_mean_fis[match(sub_res$subfacet, res$subfacet)]]
saveRDS(sub_res, "../data/sub_res_theta_proj_comp.RDS")
saveRDS(res, "../data/res_theta_proj_comp.RDS")
saveRDS(list(sfs1, sfs2, sfs3), "../data/res_sfs_comp.RDS")
```

# Plot

Read in what we ran above, only run once since it takes a while.

```{r, eval=TRUE}
res <- readRDS("../data/res_theta_proj_comp.RDS")
sub_res <- readRDS("../data/sub_res_theta_proj_comp.RDS")
sfs <- readRDS("../data/res_sfs_comp.RDS")
sfs1 <- sfs[[1]]
sfs2 <- sfs[[2]]
sfs3 <- sfs[[3]]
```

Plot
```{r}
# prepare plotting data
colours <- khroma::color("batlow")
colors <- colours(4, range=c(0.2,1))
t.sizes <- c(20, 18)
n <- 25

pd <- data.table(subfacet = c("NAM", "OutbredA", "Random"),
                 `"S"^"'"` = res$seg_sites,
                 `"Projection"` = c(sum(sfs1[-1]),
                                sum(sfs2[-1]),
                                sum(sfs3[-1])),
                 fis = res$weighted_mean_fis,
                 `"Watterson" * "'" * "s " * theta` = res$global_ws.theta)

K <- n*2 #average sample size for ws.theta, as in hohenlohe 2010. Alternative would make this into a function, then use lapply on the vector of K values
a1 <- sum(1/seq(1:(K-1))) #get a1
pd[, `"Watterson" * "'" * "s " * theta["adjusted"]` := res$seg_sites/a1]
pd[, `"Watterson" * "'" * "s " * theta["adjusted"]` :=`"Watterson" * "'" * "s " * theta["adjusted"]`/mean(`"Watterson" * "'" * "s " * theta["adjusted"]`)]
pd <- melt(pd, id.vars = c("subfacet", "fis"))
pd[,raw := value]
pd[,value := value/mean(value), by = variable]

sub_res <- as.data.table(sub_res)
sub_res[,value := seg_sites/mean(seg_sites)]

# plot
p1 <- ggplot(sub_res, aes(x = fis, y = value, group = fis)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  geom_point(data = pd,
             aes(color = variable), size = 4, position = position_dodge2(width = .005)) +
  scale_color_manual(values = colors, labels = scales::parse_format()) +
  geom_vline(xintercept = 0, linetype = "dashed")  +
  theme_bw() %+replace%
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black",
                     linewidth = rel(1)),
        axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2])) +
  ggbreak::scale_x_break(c(0.005, 0.13), ticklabels = c(.130, .135, .14)) +
  scale_x_continuous(breaks = c(-.035, -.03, -.025, -.02, -.015, -.010, -.005, 0)) +
  ylab(expression(theta * ' or S/' * mu[theta * ' or S' ])) +
  xlab(expression(italic(F)[IS])) +
  guides(color = guide_legend(title = "Method"))

ggsave("Figure5.svg", p1, height = 10, width = 15, dpi = 1000) # doesn't want to save as pdf with break
```

# Figure S6
```{r}
S6 <- ggplot(sub_res, aes(x = fis, y = seg_sites, group = fis)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  geom_point(data = pd[-grep("Watterson", variable),],
             aes(color = variable, y = raw), size = 4, position = position_dodge2(width = .005)) +
  scale_color_manual(values = colors, labels = scales::parse_format()) +
  geom_vline(xintercept = 0, linetype = "dashed")  +
  theme_bw() %+replace%
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "black",
                     linewidth = rel(1)),
        axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2])) +
 ggbreak::scale_x_break(c(0.005, 0.13), ticklabels = c(.130, .135, .14)) +
  scale_x_continuous(breaks = c(-.035, -.03, -.025, -.02, -.015, -.010, -.005, 0)) +
  ylab(bquote(italic(s))) +
  xlab(expression(italic(F)[IS])) +
  guides(color = guide_legend(title = "Method"))

ggsave("FigureS6.svg", S6, height = 10, width = 15, dpi = 1000) # doesn't want to save as pdf with break

```

### 2024-09-30 scripts/ms_supp.RMD
pre-resubmission commit

--- 2024-04-09 scripts/seg_sites_rarefaction.Rmd
@@
 threads <- 25
-allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
-n1 <- 5000 # p1 gene copy number
-n2 <- 500 # p2 gene copy number
 k <- seq(10, 100, by = 10) # numbers to rarefact to
@@
 max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
+n1 <- 2500
+n2 <- 250
 
@@
 set.seed(seed)
-nal <- length(allele_frequencies)
-miss_percent <- floor(runif(nal, 0, max_missing))/100
 
@@
 ```{r}
-# generate genotypes
-g1 <- matrix(0, nrow = nal, ncol = n1) # init
-g2 <- matrix(0, nrow = nal, ncol = n2)
-
-# add missing data
-ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
-ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
-
-# determine minor allele placements
-ma1 <- ma2 <- vector("list", length(ms1))
-for(i in 1:nal){
-  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
-  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)
-
-}
-
-# populate
-## function to replace minors
-populate <- function(g, missing, minors){
-  if(length(missing) > 0){
-    g[missing] <- NA
-    g[-missing][minors] <- 1
-  }
-  else{
-    g[minors] <- 1
-  }
-  return(g)
-}
-## do for every locus
-for(i in 1:nal){ 
-  g1[i,] <- populate(g1[i,], ms1[[i]], ma1[[i]])
-  g2[i,] <- populate(g2[i,], ms2[[i]], ma2[[i]])
-}
-
-## combine into individuals
-g1 <- g1[,1:(ncol(g1)/2)] + g1[,((ncol(g1)/2) + 1):ncol(g1)]
-g2 <- g2[,1:(ncol(g2)/2)] + g2[,((ncol(g2)/2) + 1):ncol(g2)]
+# g <- GeneArchEst::process_ms("data/ms_out.ms", 10000000)
+# 
+# g <- GeneArchEst::convert_2_to_1_column(g$x)
+# g <- t(g)
+# 
+# g1 <- g[,1:n1]
+# g2 <- g[,(n1 + 1):(n1 + n2)]
+# 
+# # derived parms
+# nal <- nrow(g1)
+# miss_percent <- floor(runif(nal, 0, max_missing))/100
+# 
+# # add missing data
+# ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
+# ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
+# 
+# # populate missing
+# 
+# ## do for every locus
+# for(i in 1:nal){ 
+#   g1[i,ms1[[i]]] <- NA
+#   g2[i,ms2[[i]]] <- NA
+# }
+# 
+# # convert ns to gene copy numbers
+# n1 <- n1*2
+# n2 <- n2*2
 
@@
 ```{r}
+set.seed(seed)
+
+gd <- readRDS("../data/g_ms_large.RDS")
+g1 <- gd$g1
+g2 <- gd$g2
+
+nal <- nrow(g1)
+
+keep_snps <- sort(sample(nal, 10000, FALSE))
+g1 <- g1[keep_snps,]
+g2 <- g2[keep_snps,]
+
+rm(gd); gc();
+
+nal <- length(keep_snps)
+allele_frequencies_g1 <- rowSums(g1, na.rm = TRUE)/(rowSums(!is.na(g1))*2)
+allele_frequencies_g1[allele_frequencies_g1 > .5] <- 1 - allele_frequencies_g1[allele_frequencies_g1 > .5]
+
+allele_frequencies_g2 <- rowSums(g2, na.rm = TRUE)/(rowSums(!is.na(g2))*2)
+allele_frequencies_g2[allele_frequencies_g2 > .5] <- 1 - allele_frequencies_g2[allele_frequencies_g2 > .5]
+
 counts <- function(g){
@@
 # Rarefact via simulations
-Note: this chunck should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunck.
+Note: this chunk should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunk.
 `q` should be passed as an argument for the iteration, as should `outfile`.
 
-```{r cluster_MCMCs, eval=FALSE}
-set.seed(seed + q) # reset seed for cluster
+```{r cluster_MCMCs}
+# set.seed(seed + q) # reset seed for cluster
 
@@
 
-# draw for each k/mcmc count combination
+# # draw for each k/mcmc count combination
 combs <- expand.grid(k, mcmcs)
 colnames(combs) <- c("k", "iters")
-
-tmcmcs <- combs$iters[q]
-tk <- combs$k[q]
-
-seg1 <- seg2 <- matrix(0, nal, tmcmcs)
-for(i in 1:tmcmcs){
-  for(j in 1:nal){
-    seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
-    seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
-  }
-}
-
-mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
-                               method = "agresti-coull")[,c("mean", "lower", "upper")]
-mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
-                               method = "agresti-coull")[,c("mean", "lower", "upper")]
-
-colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
-
-res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
-
-saveRDS(res, paste0(outfile, "_", q, ".RDS"))
+# 
+# tmcmcs <- combs$iters[q]
+# tk <- combs$k[q]
+# 
+# seg1 <- seg2 <- matrix(0, nal, tmcmcs)
+# for(i in 1:tmcmcs){
+#   for(j in 1:nal){
+#     seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
+#     seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
+#   }
+# }
+# 
+# mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
+#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
+# mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
+#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
+# 
+# colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
+# 
+# res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
+# 
+# saveRDS(res, paste0(outfile, "_", q, ".RDS"))
 
@@
 ```{r}
-res_files <- list.files("../data/", "\\.RDS", full.names = TRUE)
-res_files <- lapply(res_files, readRDS)
-ks <- unlist(purrr::map(res_files, "k"))
-n_iters <- unlist(purrr::map(res_files, "iters"))
+res_files <- list.files("../data/ms_MCMC/", "mcmc.+\\.RDS", full.names = TRUE)
+q <- gsub(".+__", "", res_files)
+q <- as.numeric(gsub("\\.RDS", "", q))
+# res_files <- pbapply::pblapply(res_files, readRDS)
+ks <- combs$k[q]
+n_iters <- combs$iters[q]
+# ks <- unlist(purrr::map(res_files, "k"))
+# n_iters <- unlist(purrr::map(res_files, "iters"))
 
@@
 ```{r Figure1}
+n1 <- n1*2
+n2 <- n2*2
 prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
-  #========plot the emperical CIs around every loci vs the predicted value==============
-  # plotting the emperical CIs vs the estiamted mean
-  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
-                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))
+  #========plot the empirical CIs around every loci vs the predicted value==============
+  # plotting the empirical CIs vs the estimated mean
+  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies_g1),
+                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies_g2))
 
-  oseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies),
-                data.frame(pop = paste0("N = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies))
+  oseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies_g1),
+                data.frame(pop = paste0("N = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies_g2))
 
@@
   seg <- rbind(pseg, oseg)
-  cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("N = ", n1/2), af = allele_frequencies),
-               cbind.data.frame(mpseg2, pop = paste0("N = ", n2/2), af = allele_frequencies))
-  colnames(cis) <- c("p", "lower", "upper", "pop", "af")
-  seg <- merge(seg, cis[,c("lower", "upper", "af", "pop")], by = c("af", "pop"))
+  cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("N = ", n1/2), af = allele_frequencies_g1, snp = 1:nal),
+               cbind.data.frame(mpseg2, pop = paste0("N = ", n2/2), af = allele_frequencies_g2, snp = 1:nal))
+  colnames(cis) <- c("p", "lower", "upper", "pop", "af", "snp")
+  seg <- merge(seg, cis[,c("lower", "upper", "pop", "snp")], by = c("snp", "pop"))
   seg$in_ci <- ifelse(seg$p >= seg$lower & seg$p <= seg$upper, TRUE, FALSE)
@@
 K10_data <- which(ks == 10 & n_iters == 10000)
+K10_data <- readRDS(res_files[[K10_data]])
 F1data_K10 <- prep_f1(pseg1 = pseg1$K_10, 
                       pseg2 = pseg2$K_10, 
-                      mpseg1 = res_files[[K10_data]]$mpseg1,
-                      mpseg2 = res_files[[K10_data]]$mpseg2,
+                      mpseg1 = K10_data$mpseg1,
+                      mpseg2 = K10_data$mpseg2,
                       n1 = n1,
@@
 K100_data <- which(ks == 100 & n_iters == 10000)
+K100_data <- readRDS(res_files[[K100_data]])
+
 F1data_K100 <- prep_f1(pseg1 = pseg1$K_100, 
                        pseg2 = pseg2$K_100, 
-                       mpseg1 = res_files[[K100_data]]$mpseg1,
-                       mpseg2 = res_files[[K100_data]]$mpseg2,
+                       mpseg1 = K100_data$mpseg1,
+                       mpseg2 = K100_data$mpseg2,
                        n1 = n1,
@@
 
-ggsave("Figure1.pdf", f1, width = 15, height = 10)
+ggsave("FigureS4.pdf", f1, width = 15, height = 10)
 tapply(F1data[F1data$source == "Mathematical", "in_ci"], F1data[F1data$source == "Mathematical", c("K", "pop")], mean) # percent in interval
@@
 K20_data <- which(ks == 20 & n_iters == 10000)
+K20_data <- readRDS(res_files[[K20_data]])
+
 F1data_K20 <- prep_f1(pseg1 = pseg1$K_20, 
                        pseg2 = pseg2$K_20, 
-                       mpseg1 = res_files[[K20_data]]$mpseg1,
-                       mpseg2 = res_files[[K20_data]]$mpseg2,
+                       mpseg1 = K20_data$mpseg1,
+                       mpseg2 = K20_data$mpseg2,
                        n1 = n1,
@@
 
-K10_data <- which(ks == 10 & n_iters == 10000)
-F2data_K10 <- prep_f2(seg1 = res_files[[K10_data]]$seg1, 
-                      seg2 = res_files[[K10_data]]$seg2, 
+F2data_K10 <- prep_f2(seg1 = K10_data$seg1, 
+                      seg2 = K10_data$seg2, 
                       vseg1 = vseg1$K_10,
@@
 
-K20_data <- which(ks == 20 & n_iters == 10000)
-F2data_K20 <- prep_f2(seg1 = res_files[[K20_data]]$seg1, 
-                       seg2 = res_files[[K20_data]]$seg2, 
+F2data_K20 <- prep_f2(seg1 = K20_data$seg1, 
+                       seg2 = K20_data$seg2, 
                        vseg1 = vseg1$K_20,
@@
         strip.background = element_blank()) +
-  xlab(bquote(italic(N[S]))) +
+  xlab(bquote(italic(S))) +
   ylab("Density")
 
-ggsave("Figure3.pdf", f2, height = 10, width = 15)
+ggsave("FigureS5.svg", f2, height = 10, width = 15)
 
 ```
-
-# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k
-Note that the labels are flipped here, this produces figure 2
-
-```{r Figure3}
-# for all of the mpseg files, figure out if the percentage of points in the CI changes
-## do for both each SNP and for Ns
-ps_list <- ns_list <- vector("list", length(res_files))
-for(i in 1:length(res_files)){
-  tk <- res_files[[i]]$k
-  titers <- res_files[[i]]$iters
-
-  ps_prep <- prep_f1(pseg1 = pseg1[[paste0("K_", tk)]],
-                     pseg2 = pseg2[[paste0("K_", tk)]], 
-                     mpseg1 = res_files[[i]]$mpseg1,
-                     mpseg2 = res_files[[i]]$mpseg2,
-                     n1 = n1,
-                     n2 = n2)
-
-  ns_prep <- prep_f2(seg1 = res_files[[i]]$seg1, 
-                     seg2 = res_files[[i]]$seg2, 
-                     vseg1 = vseg1[[paste0("K_", tk)]],
-                     vseg2 = vseg2[[paste0("K_", tk)]],
-                     F1data = ps_prep,
-                     n1 = n1,
-                     n2 = n2)
-
-  ps_prep$K <- tk
-  ps_prep$iters <- titers
-  ps_prep$sim_num <- i
-
-  ns_prep$all_mcmc_sums$K <- tk
-  ns_prep$all_mcmc_sums$iters <- titers
-  ns_prep$all_mcmc_sums$i <- i
-  ns_prep$all_mcmc_sums$ENs <- ns_prep$sum_tab$nseg[match(ns_prep$all_mcmc_sums$pop,
-                                                          ns_prep$sum_tab$pop)]
-
-  ps_list[[i]] <- ps_prep
-  ns_list[[i]] <- ns_prep$all_mcmc_sums
-}
-
-ps_all <- rbindlist(ps_list)
-
-ps_all_m <- ps_all[source == "Mathematical",]
-ps_all_s <- ps_all[source == "Simulation",]
-ps_all_merge <- merge(ps_all_m[,-c("source")], 
-                      ps_all_s[,-c("source")], 
-                      by = c("af", "pop", "snp", "sim_num", "lower", "upper", "K", "iters"), 
-                      suffixes = c(".mathematical", ".simulated"))
-ps_all_merge[,diff := p.mathematical - p.simulated]
-
-# Figure S1: gamma vs error rate
-ps_all_aggregate_for_K <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(K, iters, pop)]
-FS1 <- ggplot(ps_all_aggregate_for_K, aes(x = K, y = in_ci.mathematical*100)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(iters~pop) +
-  theme_bw() +
-  # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
-  # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
-  theme(axis.text = element_text(size = t.sizes[2] - 2),
-        axis.title = element_text(size = t.sizes[1] - 2),
-        strip.text = element_text(size = t.sizes[1] - 2),
-        strip.background = element_blank(),
-        legend.title = element_text(size = t.sizes[1] - 2),
-        legend.text = element_text(size = t.sizes[2] - 2)) +
-  xlab(parse(text = "gamma")) +
-  ylab("% estimates in CI")
-
-save_plot("FigureS1.pdf", FS1, base_height = 10, base_width = 15)
-
-# # Figure S2: maf vs error rate
-# ps_all_aggregate_for_af <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(af, iters, pop)]
-# FS2 <- ggplot(ps_all_aggregate_for_af, aes(x = af, y = in_ci.mathematical*100)) +
-#   geom_point() +
-#   geom_smooth(method = "lm", se = FALSE) +
-#   facet_grid(iters~pop) +
-#   theme_bw() +
-#   # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
-#   # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
-#   theme(axis.text = element_text(size = t.sizes[2]),
-#         axis.title = element_text(size = t.sizes[1]),
-#         strip.text = element_text(size = t.sizes[1]),
-#         strip.background = element_blank(),
-#         legend.title = element_text(size = t.sizes[1]),
-#         legend.text = element_text(size = t.sizes[2])) +
-#   xlab("Minor Allele Frequency") +
-#   ylab("% estimates in CI")
-
-ps_all_merge[,in_ci.mathematical := ifelse(in_ci.mathematical, "Yes", "No")]
-ps_all_merge[,in_ci.mathematical := factor(in_ci.mathematical, c("Yes", "No"))]
-f3a <- ggplot(ps_all_merge, aes(x = K, y = diff, color = in_ci.mathematical, group = in_ci.mathematical)) +
-  geom_point(position = position_dodge(4)) +
-  theme_bw() +
-  facet_grid(~pop) +
-  khroma::scale_color_highcontrast() +
-  scale_x_continuous(breaks = k) +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
-        strip.background = element_blank(),
-        legend.title = element_text(size = t.sizes[1]),
-        legend.text = element_text(size = t.sizes[2])) +
-  xlab(parse(text = "gamma")) +
-  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")"))) +
-  guides(color = guide_legend(title = "In 95% CI"))
-
-f3b <- ggplot(mapping = aes(x = af, y = diff, color = in_ci.mathematical)) +
-  geom_point(data = ps_all_merge[in_ci.mathematical == "Yes",]) +
-  geom_point(data = ps_all_merge[in_ci.mathematical == "No",]) +
-  theme_bw() +
-  facet_grid(~pop) +
-  khroma::scale_color_highcontrast() +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_blank(),
-        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
-        strip.background = element_blank(),
-        legend.title = element_text(size = t.sizes[1]),
-        legend.text = element_text(size = t.sizes[2])) +
-  xlab("Minor Allele Frequency") +
-  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")")))
-
-f3c <- ggplot(ps_all_merge, aes(x = iters, y = diff, color = in_ci.mathematical, group = in_ci.mathematical)) +
-  geom_point(position = position_dodge(400)) +
-  theme_bw() +
-  facet_grid(~pop) +
-  khroma::scale_color_highcontrast() +
-  scale_x_continuous(breaks = mcmcs) +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        strip.text = element_blank(),
-        axis.text.x = element_text(size = t.sizes[2], 
-                                   angle = 90, vjust = .5),
-        strip.background = element_blank(),
-        legend.title = element_text(size = t.sizes[1]),
-        legend.text = element_text(size = t.sizes[2])) +
-  xlab("Number of Simulated Draws") +
-  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")")))
-
-f3l <- ggpubr::get_legend(f3a)
-f3 <- plot_grid(f3a + guides(color = "none"), 
-          f3b + guides(color = "none"), 
-          f3c + guides(color = "none"),
-          ncol = 1, 
-          align = "bl", 
-          axis = "bl")
-
-f3 <- plot_grid(f3, f3l, nrow = 1, rel_widths = c(1, .2))
-
-save_plot("Figure2.pdf", f3, base_height = 11, base_width = 15)
-```
-# Figure 4 Nseg vs parameters
-
-```{r Figure4}
-ns_all <- rbindlist(ns_list)
-
-ns_all$diff <- ns_all$ENs - ns_all$nseg
-
-f4a <- ggplot(ns_all[iters == 10000,], aes(x = K, y = nseg)) +
-  geom_errorbar(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper")]),
-            aes(x = K + 3.7, ymin = lower, ymax = upper), color = "#DDAA33", size = 1,
-            inherit.aes = FALSE, width = 2) +
-  geom_point(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper", "ENs")]),
-             aes(x = K + 3.7, y = ENs), color = "#DDAA33", size = 7, shape = 45,
-            inherit.aes = FALSE) +
-  stat_ydensity(aes(group = K), draw_quantiles = c(.025, .5, .975), scale = "width", bw = .4,
-                fill = NA, width = 6) +
-  facet_wrap(~pop) +
-  scale_y_continuous(breaks = seq(40, 100, by = 10), limits = c(40, 100)) +
-  theme_bw() + 
-  scale_x_continuous(breaks = k) +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        legend.title = element_text(size = t.sizes[1]),
-        strip.text = element_text(size = t.sizes[1]),
-        strip.background = element_blank(),
-        legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(N[S]))) +
-  xlab(bquote(gamma))
-
-f4b <- ggplot(ns_all[K == 50,], aes(x = iters, y = nseg)) +
-  geom_errorbar(data = unique(ns_all[K == 50,c("pop", "K", "iters", "i", "lower", "upper")]),
-            aes(x = iters + 370, ymin = lower, ymax = upper), color = "#DDAA33", size = 1,
-            inherit.aes = FALSE, width = 200) +
-  geom_point(data = unique(ns_all[K == 50, c("pop", "K", "iters", "i", "lower", "upper", "ENs")]),
-             aes(x = iters + 370, y = ENs), color = "#DDAA33", size = 7, shape = 45,
-            inherit.aes = FALSE) +
-  stat_ydensity(aes(group = iters), draw_quantiles = c(.025, .5, .975), scale = "width", bw = .4,
-                fill = NA, width = 600) +
-  facet_wrap(~pop) +
-  theme_bw() + 
-  scale_y_continuous(breaks = seq(40, 100, by = 5), limits = c(80, 100)) +
-  scale_x_continuous(breaks = mcmcs) +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        legend.title = element_text(size = t.sizes[1]),
-        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
-        strip.text = element_blank(),
-        strip.background = element_blank(),
-        legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(N[S]))) +
-  xlab("Number of Simulated Draws")
-
-f4 <- plot_grid(f4a, f4b, ncol = 1, axis = "l", align = "l")
-save_plot("Figure4.pdf", f4, base_height = 11, base_width = 15)
-
-# S2
-ns_all_agg <- ns_all[,lapply(.SD,mean), .SDcols = "nseg", by = c("ENs", "pop", "K", "iters", "i")]
-levs <- paste0("Sims = ", sort(unique(ns_all_agg$iters)))
-ns_all_agg$iters <- factor(paste0("Sims = ", ns_all_agg$iters), levels = levs)
-ns_all_agg$pop <- gsub("n", "N", ns_all_agg$pop)
-ns_all_agg$stat <- "Mean"
-
-ns_all_agg_med <- ns_all[,lapply(.SD,median), .SDcols = "nseg", by = c("ENs", "pop", "K", "iters", "i")]
-levs <- paste0("Sims = ", sort(unique(ns_all_agg_med$iters)))
-ns_all_agg_med$iters <- factor(paste0("Sims = ", ns_all_agg_med$iters), levels = levs)
-ns_all_agg_med$pop <- gsub("n", "N", ns_all_agg_med$pop)
-ns_all_agg_med$stat <- "Median"
-
-ns_all_agg <- rbind(ns_all_agg, ns_all_agg_med)
-FS2 <- ggplot(ns_all_agg, aes(x = K, y = ENs - nseg, color = pop)) +
-  geom_point() +
-  facet_grid(stat~iters) +
-  khroma::scale_color_highcontrast() +
-  geom_hline(yintercept = 0) +
-  theme_bw() +
-  theme(axis.text = element_text(size = t.sizes[2]),
-        axis.title = element_text(size = t.sizes[1]),
-        legend.title = element_text(size = t.sizes[1]),
-        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
-        strip.background = element_blank(),
-        strip.text = element_text(size = t.sizes[1]),
-        legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(E(N[S])) * " - " * italic(N[S,observed]))) +
-  xlab(bquote(gamma)) +
-  guides(color = guide_legend(title = "Population"))
-
-save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 15)
-
-```
-
-```{r}
-# show difference in HWE divergence with different combinations of 5 minor alleles in 100 genotypes
-exd <- rbind(c(rep(0, 45), rep(1, 5)),
-             c(rep(0, 46), 2, rep(1, 3)),
-             c(rep(0, 47), rep(2, 2), 1))
-exds <- import.snpR.data(as.data.frame(exd), snp.meta = data.frame(snp = 1:3), mDat = -9)
-exds <- calc_hwe(exds)
-get.snpR.stats(exds, stats = "hwe")
-
-exc <- counts(exd)
-rarefact(exc, 10)
-```

### 2024-09-30 scripts/seg_sites_rarefaction.Rmd
pre-resubmission commit

@@
 threads <- 25
-allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
+allele_frequencies <- seq(.001, .5, by = .001) # allele frequencies
 n1 <- 5000 # p1 gene copy number
@@
         strip.background = element_blank()) +
-  xlab(bquote(italic(N[S]))) +
+  xlab(bquote(italic(S))) +
   ylab("Density")
@@
   facet_wrap(~pop) +
-  scale_y_continuous(breaks = seq(40, 100, by = 10), limits = c(40, 100)) +
+  scale_y_continuous(breaks = seq(430, 500, by = 10), limits = c(430, 500)) +
   theme_bw() + 
@@
         legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(N[S]))) +
+  ylab(bquote(italic(S))) +
   xlab(bquote(gamma))
@@
   theme_bw() + 
-  scale_y_continuous(breaks = seq(40, 100, by = 5), limits = c(80, 100)) +
+  scale_y_continuous(breaks = seq(480, 500, by = 5), limits = c(480, 500)) +
   scale_x_continuous(breaks = mcmcs) +
@@
         legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(N[S]))) +
+  ylab(bquote(italic(S))) +
   xlab("Number of Simulated Draws")
@@
         legend.text = element_text(size = t.sizes[2])) +
-  ylab(bquote(italic(E(N[S])) * " - " * italic(N[S,observed]))) +
+  ylab(bquote(italic(S^"'") * " - " * italic(S))) +
   xlab(bquote(gamma)) +
@@
   theme_bw() +
-  ylab(bquote(italic(N[s]))) +
+  ylab(bquote(italic(S))) +
   xlab("Population") +

### 2024-09-30 scripts/phase_monarchs.R

@@ -1,0 +1,185 @@
+library(snpR); library(ggplot2)
+d <- readRDS("../ConGen/microConGen_RW/Data/monarch_nomaf.RDS")
+
+chrs <- table(snp.meta(d)$group)
+bad_chrs <- names(chrs)[which(chrs == 1)]
+bad_snps <- which(snp.meta(d)$group %in% bad_chrs)
+d <- import.snpR.data(genotypes(d)[-bad_snps,], snp.meta(d)[-bad_snps,], sample.meta(d))
+
+format_snps(d, "vcf", chr = "group", outfile = "data/monarchs.vcf")
+cmd <- "java -Xmx4g -jar ~/bin/beagle.22Jul22.46e.jar gt=data/monarchs.vcf out=data/monarchs_phased"
+system(cmd)
+system("gunzip data/monarchs_phased.vcf.gz")
+
+
+dv <- data.table::fread("data/monarchs_phased.vcf")
+pops <- colnames(dv)[-c(1:9)]
+pops <- substr(pops, 1, 3)
+
+
+meta <- dv[,1:2]
+dv <- dv[,-c(1:9), with = FALSE]
+
+
+genotypes <- list(nam = dv[,which(pops == "NAM"), with = FALSE],
+                  qld = dv[,which(pops == "QLD"), with = FALSE])
+genotypes <- lapply(genotypes, function(x){
+  x1 <- matrix(as.numeric(substr(as.matrix(x), 1, 1)), nrow = nrow(x))
+  x2 <- matrix(as.numeric(substr(as.matrix(x), 3, 3)), nrow = nrow(x))
+  return(cbind(x1, x2)[,order(c(1:ncol(x1), 1:ncol(x2)))])
+})
+
+effects <- rep(0, nrow(dv))
+effects <- cbind(effects, effects)
+
+
+target_size <- 50
+migration <- matrix(.5, 2, 2)
+chr_lengths <- tapply(meta$POS, meta$`#CHROM`, max)
+colnames(meta) <- c("chr", "position")
+set.seed(123343)
+tno <- simulate_populations(genotypes, meta = meta, effects = effects, h = .5, gens = 1, chr_length = chr_lengths,
+                            growth_function = function(n) BL_growth(n, 2),
+                            var_theta = 0,
+                            K_thin_post_surv = target_size,
+                            thin = FALSE,
+                            thin_fixed = FALSE, 
+                            sampling_point = "offspring",
+                            starting_surv_opt = c(0, 0),
+                            migration = migration,
[…]

### 2024-09-30 scripts/watersons_proj_exploration.R

@@ -1,0 +1,143 @@
+library(ggplot2); library(snpR)
+
+d <- GeneArchEst::process_ms("../genomic_prediction_accuracy/data/theta4k_1000_10_rho40k.txt", 10000000)
+
+keeps <- sample(nrow(d$meta), 10000, FALSE)
+d$x <- d$x[keeps,]
+d$meta <- d$meta[keeps,]
+genotypes <- list(d$x[,1:250])
+
+meta <- d$meta
+effects <- rep(0, nrow(meta))
+
+effects <- cbind(effects, effects, effects)
+inbreeding <- c(10, -10, 0)
+
+target_size <- 50
+set.seed(123343)
+tno <- simulate_populations(genotypes[rep(1, 3)], meta = meta, effects = effects, h = .5, gens = 10, chr_length = 10000000,
+                            growth_function = function(n) BL_growth(n, 2), # use negative binomial instead
+                            var_theta = 0,
+                            K_thin_post_surv = target_size,
+                            thin = FALSE,
+                            thin_fixed = TRUE, 
+                            sampling_point = "offspring",
+                            inbreeding = inbreeding,
+                            starting_surv_opt = c(0, 0, 0),
+                            survival_function = function(phenotypes, ...) rep(1, length(phenotypes)),
+                            selection_shift_function = function(opt, ...) opt,
+                            verbose = TRUE)
+
+
+genos <- cbind(tno$genotypes[[1]], tno$genotypes[[2]], tno$genotypes[[3]])
+genos <- GeneArchEst::convert_2_to_1_column(genos)
+genos <- t(genos)
+d <- import.snpR.data(as.data.frame(genos), snp.meta = tno$meta, 
+                      sample.meta = data.frame(pop = c(rep("Inbred", ncol(tno$genotypes[[1]])/2),
+                                                       rep("Outbred", ncol(tno$genotypes[[2]])/2),
+                                                       rep("Random", ncol(tno$genotypes[[3]])/2))))
+
+
+
+set.seed (3218434)
+d <- filter_snps(d, non_poly = TRUE)
+d <- calc_fis(d, "pop")
+get.snpR.stats(d, "pop", "fis")$weighted.means
+
+
+n <- 25
+
+d <- calc_seg_sites(d, "pop", g = n)
[…]

### 2024-10-01 scripts/seg_sites_rarefaction.Rmd
minor fixes, added line numbers to response

@@
 ggsave("Figure1.pdf", f1, width = 15, height = 10)
-tapply(F1data[F1data$source == "Mathematical", "in_ci"], F1data[F1data$source == "Mathematical", c("K", "pop")], mean) # percent in interval
+tapply(F1data[F1data$source == "Mathematical", "in_ci"], F1data[F1data$source == "Mathematical", c("K", "pop")], function(x) mean(ifelse(x == "Yes", TRUE, FALSE))) # percent in interval
+mean(ifelse(F1data[F1data$source == "Mathematical", "in_ci"] == "Yes", TRUE, FALSE))
 ```
